### 2022-05-18 Manuscript/inter_spike_manuscript.tex
added manuscript

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
\documentclass[smallcondensed]{svjour3}    % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn

%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\setlength{\marginparwidth}{2cm}
\usepackage{changes}
\usepackage{bbm}
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{mathptmx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{statmath}
\usepackage[numbers]{natbib}
\usepackage{hyperref}
\usepackage{enumitem}
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
\journalname{Nonlinear Dynamics}
%
\begin{document}

\title{Recurrence spike spectra}

\titlerunning{Recurrence spike spectra}        % if too long for running head

\author{K. Hauke Kraemer \and
		Frank Hellmann \and
        J\"urgen Kurths \and 
        Norbert Marwan
}

\institute{K. Hauke Kraemer \and J\"urgen Kurths \and Norbert Marwan \and Frank Hellmann \at Potsdam Institute for Climate Impact Research, Member of the Leibniz Association, 14473 Potsdam, 
Germany \email{hkraemer@pik-potsdam.de}\and
           K. Hauke Kraemer \and J\"urgen Kurths \at Institute of Physics and Astronomy, University of Potsdam, 14476 Potsdam, Germany \and
		   Norbert Marwan \at Institute of Geosciences, University of Potsdam, 14476 Potsdam, Germany \and
		   J\"urgen Kurths \at Institute of Physics, Humboldt Universit\"at zu Berlin, 12489 Berlin, Germany
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
A novel kind of power spectrum is constructed, the \textit{inter spike spectrum}, which transforms spike-train-like signals into their frequency domain. This method 
clearly shows the apparent cycles in the data and overcomes the problems when using the obvious idea of Fourier-transforming it. We invent this instructive approach with the 
idea of transforming the $\tau$-recurrence rate of a recurrence plot (RP), which often has a spiky appearance. The $\tau$-recurrence rate is the density of recurrence points 
along diagonals of the RP, which are parallel to the main diagonal with a distance of $\tau$. In this context the inter spike spectrum can be interpreted as a nonlinear power 
spectrum of a potentially high dimensional system which constitutes the RP. The proposed measure is simple to compute, robust to noise and is able to detect bifurcations 
inducing regular-regular, regular-chaos as well as chaos-chaos transitions.

\keywords{Bifurcations \and Recurrence Analysis \and Decomposition \and Frequency Analysis}
\PACS{05.45.Tp, 05.90.+m, 89.90.+n, 02.70.Uu, 05.10.Ln, 05.45.-a, 05.45.Ac}
\end{abstract}

\section{Introduction}\label{sec_tau_rr_intro}

Recurrence Plots (RPs) provide a vivid representation of complex dynamics stemming from potentially high dimensional systems, Eq.~\eqref{eq_rp_definition}. 
\begin{equation}\label{eq_rp_definition}
R_{i,j}(\varepsilon) = \Theta\left(\varepsilon - D_{i,j}\right) 
= \Theta\left(\varepsilon - \| \vec{x}_i - \vec{x}_j\|\right), \qquad \vec{x} \in \mathbb{R}^d, \quad i,j \in [1,\ldots, N].
\end{equation}
The simple idea to 
track recurring states of the $d$-dimensional trajectory $\vec{x}_i$ of the system under study not only allows for a beneficial visualization of the dynamics, but also for its 
quantification, using certain structures in the RP, such as diagonal or vertical lines \cite{marwan2007}.

Some of these recurrence quantification measures, the entropy of diagonal lines and the entropy of 
recurrence times, can be related to the Kolmogorov-Sinai entropy \cite{march2005,baptista2010}. However, these quantifiers have a free parameter, the minimal considered line length, and 
are usually biased, due to the finite size of the RP and thickened diagonal lines, which needs to be corrected \cite{Kraemer2019}. Moreover, the mentioned statistics cannot account for 
changing regular (non-chaotic) dynamics, such as period-doubling bifurcations.
A rather simple idea is to look at the $\tau$-recurrence rate of the RP ($\tau$-RR, Eq.~\ref{eq_tau_rr}) \cite{marwan2002pla,Zbilut2008}.
This is the density of recurrence points along the diagonals of the recurrence matrix, as a function of the distance $\tau$ (sampling units) to the main diagonal:
\begin{equation}\label{eq_tau_rr}
\tau\text{-RR}(\varepsilon) = RR(\tau, \varepsilon) = \frac{1}{N-\tau} \sum_{i=1}^{N-\tau	} R_{i,i+\tau}.
\end{equation}
$\tau$-RR serves as an estimator for the probability that the system recurs after time $\tau \Delta t$, with $\Delta t$ being the sampling time of the trajectory 
$\vec{x}_i = \vec{x}(\Delta t \cdot i),\, i=1,\ldots,N$. 
\citet{Zbilut2008} pointed out that $\tau$-RR could be used as a plugin value for the auto-correlation function $C(\tau)$ and, hence, via the Wiener-Khinchim theorem a 
``generalized'' powerspectrum can be obtained. This is reasonable, since the average distances for a given lag $\tau$ 
\begin{equation}
\overline{D}(\tau) = \frac{1}{N-\tau}\sum_{i=1}^{N-\tau} D_{i, i+\tau}
\end{equation}
can be directly read from the distance matrix $\mathbf{D}$ and is also preserved in its thresholded version $\tau$-RR. There are clear advantages for a recurrence-derived 
powerspectrum, i.e., Fourier transforming (FT) $\tau$-RR (Fig.~\ref{fig_tau_rr_spectrum_example}D), instead 
of $C(\tau)$ (Fig.~\ref{fig_tau_rr_spectrum_example}C): There are no assumptions for stationarity or sampling, when constructing a RP.
Furthermore, the correlation structures of higher dimensional spaces can be resolved in the recurrence-derived Fourier-spectrum.\\

\begin{figure}
 \centering
 \includegraphics[width=0.9\textwidth]{./figures/fig_tau_rr_spectrum_example}
 \caption{Schematic illustration of a $\tau$-recurrence rate based spectrum. \textbf{A} $x$-component time series of the Lorenz63-System (Eq.~\eqref{eq_model_Lorenz63}) and 
 \textbf{B} its corresponding Fourier power spectrum. 
 \textbf{C} Reconstructed state space portrait from the time series shown in \textbf{A} using PECUZAL time-delay embedding \cite{Kraemer2021}. 
 \textbf{D} Subset of the recurrence plot and the corresponding $\tau$-recurrence rate obtained from the state space trajectory in \textbf{C}. The shaded interval in the time series in \textbf{A} corresponds to the shown subset. 
 \textbf{E} Fourier Power spectrum obtained from the $\tau$-recurrence (subset shown in panel \textbf{D}) \cite{Zbilut2008}.
 }\label{fig_tau_rr_spectrum_example}
\end{figure}

However, there is a drawback to this approach. Whenever $\tau$-RR is a spike-train-like signal, which it is in most cases (see Fig.~\ref{fig_tau_rr_spectrum_example}) especially for 
map-data (low-resolution data), a FT of such a signal leads to a spike-train-like image in the frequency domain (e.g. \cite{Schild1982,Cordoba1989}, see 
Fig.~\ref{fig_tau_rr_spectrum_example}D). Thus, it is not intuitive how to extract meaningful information about dominant frequencies of the systems' state space trajectory.

For clarification, consider the signal we would like to analyze 
(e.g. the $\tau$-RR of a system) to be a Dirac comb (DC) with inter-spike period $T_\text{is}$: 
\begin{equation}
\text{DC}_{\text{is}}(t) = \sum_{k=-\infty}^{\infty} \delta(t-kT_\text{is}),
\label{eq_dirac_comb}
\end{equation}
i.e., a series of Dirac delta functions for a period $T_\text{is}$. There is only one single period -- $T_\text{is}$ -- in this signal (Fig.~\ref{fig_tau_rr_dirac_comb}A, D), so 
in principle we would strive for a single peak in the frequency domain of this signal at a frequency $f=1/T_\text{is}$. Surprisingly the Fourier spectrum does not meet this expectation 
and instead of a single frequency, there are exceptionally many frequencies excited (Fig.~\ref{fig_tau_rr_dirac_comb}B, E). This is, because the Fourier components add constructively 
for every frequency $1/T_\text{is}$ and therefore $\text{DC}_{\text{is}}(t)$ coincides with its own Fourier transform up to a factor $1/T_\text{is}$.

\begin{figure}[h]
 \centering
 \includegraphics[width=0.9\textwidth]{./figures/fig_tau_rr_dirac_comb}
 \caption{The transformation of a Dirac comb (series of Dirac delta functions) with a single inter-spike period $T_{\text{is}}=100$ ($\widehat{=}f=0.01$) 
 into the frequency domain. 
 \textbf{A} Dirac Comb (DC) with equal amplitudes and
 \textbf{B} its FFT-based powerspectrum.
 \textbf{C} Proposed inter spike spectrum of the signal in \textbf{A} showing a single frequency, which corresponds to the inter-spike period $T_{\text{is}}$ ($f=0.01$).
 \textbf{D} DC with randomly chosen amplitudes and same $T_{\text{is}}$ as in \textbf{A}, and
 \textbf{E} its FFT-based powerspectrum. 
 \textbf{F} Proposed inter spike spectrum of the signal in \textbf{D} showing a single frequency, which corresponds to the expected inter-spike period $T_{\text{is}}$ ($f=0.01$).  
}
\label{fig_tau_rr_dirac_comb}
\end{figure}

In this article we propose a new way of transforming a spike-train-like signal into its frequency domain. This novel \textit{inter spike spectrum} does not show resonance 
behavior of the signal's inherent inter-spike frequencies (Fig.~\ref{fig_tau_rr_dirac_comb}C, F). Section \ref{sec_tau_rr_method} explains the idea, which can be used to decompose 
any arbitrary signal and is not restricted to the $\tau$-RR, which we exemplify here. However, the more spiky the signal is, the more outstanding our new approach is compared to a 
FT. In Section \ref{sec_tau_rr_application} we exemplify its use when transforming the $\tau$-RR of a system under study. In this case, the inter spike spectrum can unravel 
characteristic time scales of high dimensional systems, which is not possible when using a FT.

\section{Method}\label{sec_tau_rr_method}

The idea is as simple as it is obvious. The signal, which we would like to transform -- we focus on $\tau$-RR in this article, but this can be applied to any sort of signal -- 
is decomposed into a set of appropriate basis functions. Instead of using trigonometric functions, as it is the idea in the Fourier decomposition, we use Dirac combs (DC) with 
different inter-spike periods as basis functions, Eq.~\eqref{eq_dirac_comb}. Let $s(t_i)$ be the normalized signal we want to transform with length $N$ and 
$t_i=i\cdot \Delta t,~i=1,\ldots,N$, where $\Delta t$ denotes the sampling time and $s(t_i) \in [0,\ 1]\ \forall\ i$. In the following we label this time series as a $(1\times N)$-dimensional vector $\bf{s}$. 
First, $N$ different DC's of length $N$ are constructed with inter-spike periods $T_\text{is} \in [1,\ldots,N]$. Second, in order to account for possible phase shifts of 
these basis functions occurring 
in $\bf{s}$, each of these $N$ different DC's also need to be shifted one step further $T_\text{is}-1$ times. This leaves us with a total number of $M = \sum_{i=1}^{N}i$ 
basis functions which we can arrange as rows of a $(M\times N)$-sized matrix $\bfX$
(Fig.~\ref{fig_tau_rr_basis_functions} illustrates the described procedure)
\begin{align}
\bfX_{i,j} &= \sum_{k=0}^N\delta\left(j-1-kT(i)-i+T(i) \right),\\
T(i)&= n \quad, \forall n:~ \frac{n(n-1)}{2} + 1 \leq i <  \frac{n(n+1)}{2} + 1,\quad n \in \mathbb{N}_+.
\label{eq_basis_matrix}
\end{align}
Due to the shifting of each of the basis functions of inter-spike period $T_\text{is}$, 
$\bfX$ is not linear independent anymore. Thus, an under-determined linear system
\begin{equation}
\bf{X}^T\bf{\beta}=\bf{s}
\label{eq_linear_system}
\end{equation}  
has to be solved for $\bf{\beta}$, a $(M\times 1)$-sized vector carrying the loadings we are interested in. Along a variety of algorithms, which can solve this problem, we are 
particularly interested in those solutions, which promote sparsity in $\bf{\beta}$, since our goal is to decompose the signal $\bfs$ into a minimal number of basis 
functions (for an excellent overview of the topic we refer to \citet{Brunton2019}). Therefore, we here choose the \textit{least absolute shrinkage and selection operator} 
(LASSO regression) \cite{Tibshirani1996}. Finally, we group loadings which correspond to basis functions having the same period $T_\text{is}$ into $\bf{\beta}_f$ and obtain the 
\textit{inter spike spectrum} by simply plotting $\bf{\beta}_f$ as a function of the frequency $f=T_\text{is}^{-1}$ (Fig.~\ref{fig_tau_rr_dirac_comb}C, F).

\begin{figure}
  \begin{minipage}[b]{0.3\textwidth}
    \vspace{0pt}
    \includegraphics[width=0.95\textwidth]{./figures/fig_tau_rr_basis_functions}
  \end{minipage}\hfill
  \begin{minipage}[b]{0.7\textwidth}
    \vspace{0pt}
    \caption{Example for a set of basis functions for an input signal of length $N=5$, aligned in the matrix $\bfX$.} \label{fig_tau_rr_basis_functions}
  \end{minipage}
\end{figure}

\section{Application}\label{sec_tau_rr_application}

We exemplify the use of the inter spike spectrum in combination with the $\tau$-RR as outlined in 
Section \ref{sec_tau_rr_intro} on several interesting research questions. The procedure is the following:
\begin{itemize}[noitemsep]
\item[(1)] Compute a RP of the trajectory of the system, Eq.~\eqref{eq_rp_definition}. 
\item[(2)] Compute the $\tau$-RR of that RP, Eq.~\eqref{eq_tau_rr}.
\item[(3)] Transform the $\tau$-RR into the proposed inter spike spectrum, see Section \ref{sec_tau_rr_method}.
\end{itemize}

\subsection{Period estimation for different dynamics in the R\"ossler system}
First, we consider the R\"ossler system (Eq.~\eqref{eq_model_roessler} in Appendix \ref{sec_models_roessler}) 
in three different dynamical setups. We use the proposed inter spike spectrum to
identify the type of dynamics.
We set the parameters $b=2$, $c=4$ and analyze period-2 limit cycle dynamics ($a=0.36$, Fig.~\ref{fig_tau_rr_example_roessler}A, D, G, J), 
period-3 limit cycle dynamics ($a=0.41$, Fig.~\ref{fig_tau_rr_example_roessler}B, E, H, K) and chaotic dynamics ($a=0.428$, Fig.~\ref{fig_tau_rr_example_roessler}C, F, I, L).

\begin{figure}
 \centering
 \includegraphics[width=0.9\textwidth]{./figures/fig_tau_rr_example_roessler}
 \caption{Inter spike spectra of the $\tau$-RR of the R\"ossler system in three different dynamical regimes with parameters $b=2$, $c=4$. 
 Trajectory of the system in a \textbf{A} period-2 (parameter $a=0.36$), \textbf{B} in a period-3 (parameter $a=0.41$) and 
 \textbf{C} in a chaotic regime (parameter $a=0.428$). 
 \textbf{D, E, F} The corresponding RPs, obtained by using a recurrence threshold corresponding to a 10\% global 
 recurrence rate for D \& E and 5\% for F. 
  \textbf{G, H, I} $\tau$-RR's of the shown RPs. 
  \textbf{J, K, L} The proposed inter spike spectra of the $\tau$-RR's shown in panels G, H, I. Spectra were obtained with a regularization threshold corresponding to 
  $\rho=0.95$ accordance of $\tau$-RR's and re-composed signals. The distance ratio of the peaks reflect the limit cycle dynamic.  
  \textbf{M, N, P} Fourier power spectra of the $\tau$-RR's shown in panels G, H, I.
 }
\label{fig_tau_rr_example_roessler}
\end{figure}

The inter spike spectra unravel the specific dynamics, which are also apparent in the state space portraits (Fig.~\ref{fig_tau_rr_example_roessler}A, B, C) and in the 
$\tau$-RRs (Fig.~\ref{fig_tau_rr_example_roessler}G, H, I). The proposed idea is also 
robust to noise (see Fig.~\ref{fig_tau_rr_example_roessler_noise} in the Appendix). This is because the peaks of the $\tau$-RR are insensitive to noise. While the peak shape does 
change in the presence of noise, its position does not, and this is what the inter spike spectrum encrypts after all.

\subsection{Bifurcations in the Logistic map}

We consider the Logistic map $x_{n+1}=r\cdot x_n \left( 1-x_n \right)$ for changing control parameter $r$. We vary $r$ from $r=3.4$ to $r=4$ in steps of $0.001$. For 
each setting of $r$ 
\begin{itemize}[noitemsep]
\item[(1)] a time series of length $N=201$ is computed with a random initial condition $u_0 \in [0,\ 1]$, neglecting the first $1,000$ samples as transients,
\item[(2)] $100$ iterative Amplitude Adjusted Fourier Transform (iAAFT) surrogates \cite{Schreiber1996,Schreiber2000} are computed,
\item[(3)] the time series and its iAAFT surrogates are embedded in a 2-dimensional state space using a time delay of unity,
\item[(4)] from the 2-dimensional trajectories RPs, Eq.~\eqref{eq_rp_definition}, are computed under a threshold $\varepsilon=0.05$,
\item[(5)] $\tau$-RR, Eq.~\eqref{eq_tau_rr}, is computed from the RP of the signal and from the RPs of the surrogates,
\item[(6)] inter spike spectra are obtained from $\tau$-RR of the signal and from the $\tau$-RRs of the surrogates, see Section \ref{sec_tau_rr_method}, and finally,
\item[(7)] from the distribution of the surrogate inter spike spectra the $95^\text{th}$ percentile is computed. The peaks of the inter spike spectrum of the signal which exceed 
this percentile are counted. 
\end{itemize}
In this example, the according null hypothesis is that the data stems from a process which yields the same auto-correlation, hence the same Fourier powerspectrum, and the same 
amplitude distribution. We consider the number of significant peaks in the inter spike spectrum with
respect to the control parameter in order to distinguish the corresponding dynamics (Fig.~\ref{fig_tau_rr_logistic}C).
A correlation with the positive Lyapunov exponent (Fig.~\ref{fig_tau_rr_logistic}A) is discernible 
($\rho_{\text{Pearson}}=0.88$). Moreover, this analysis can tackle period-doubling, since it ``measures'' the dominant cycles via the inter spike spectrum.

A less computationally intensive approach is to compute surrogates for the $\tau$-RR analytically, rather than computing a RP and its $\tau$-RR for each iAAFT surrogate of the 
time series. This translates into a null hypothesis 
that the $\tau$-RR and its corresponding inter spike spectrum stems from a RP of a random signal. In this case the probability of finding a black point in the RP can be obtained 
from a binomial distribution with probability parameter $p$ set to the recurrence rate of the RP of the signal. This way $100$ surrogate $\tau$-RRs are computed in step (5). 
The results are similar to the ones obtained from the iAAFT surrogates and nearly as good (Fig.~\ref{fig_tau_rr_logistic}B). The first period doubling at 
$r \approx 3.458$ cannot be detected, while the correlation to the positive Lyapunov exponent is only slightly worse ($\rho_{\text{Pearson}}=0.84$).

The described procedure does work well for map data, because most often the $\tau$-RR for those kind of data reveals a ``spiky enough'' nature. 
On the contrary, highly sampled (flow-) data often yield not as 
spiky $\tau$-RR's and, thus, the number of significant peaks in the inter spike spectrum may not be sensitive enough to detect period-doubling bifurcations. Nevertheless the 
according inter spike spectra is still revealing important information (Fig.~\ref{fig_tau_rr_example_roessler}) and practitioners can design appropriate quantifying statistics based 
on these spectra, which suit the 
research task.

\begin{figure}
 \centering
 \includegraphics[width=\textwidth]{./figures/fig_tau_rr_logistic}
 \caption{\textbf{A} Lyapunov exponent of the Logistic map as a function of the control parameter $r$. 
 \textbf{B} Number of significant peaks ($\alpha=0.05$) in the inter spike spectrum of the $\tau$-RR, random surrogates. 
 \textbf{C} Same as \textbf{B}, but for iterative Amplitude Adjusted Fourier Transform (iAAFT) surrogates \cite{Schreiber1996,Schreiber2000}.
}
\label{fig_tau_rr_logistic}
\end{figure}

\subsection{Evolutionary spike spectrogram of CENOGRID isotope time series}

%To demonstrate XXXXXX we applied it to the recently published CENOzoic Global Reference
%benthic foraminifer carbon and oxygen Isotope Dataset (CENOGRID) \cite{Westerhold2020}. 
%The temperature-dependent fractionation of carbon and oxygen isotopes in benthic foraminifera
%is an important means to reconstruct past global temperatures and environmental conditions.
%Moreover, the Cenozoic is interesting, because it provides an analogue of future greenhouse 
%climate and how and which regime shifts in large-scale atmospheric and ocean circulation 
%can be expected in the future warming climate. The dataset consists of a detrended $\delta^{18}$O and a detrended $\delta^{13}$C isotope record with a total length
%of $N=13,421$ samples and a sampling period of $\Delta t= 5,000 \text{yrs}$ (Figure \ref{fig_time_series_cenogrid} in Appendix \ref{sec_appendix_cenogrid}). Here we made predictions 
%on the $\delta^{13}$C isotope record.
%
%We again utilize the Cenozoic Global Reference benthic foraminifer carbon and oxygen Isotope Dataset (CENOGRID) \cite{Westerhold2020}. The time series 
%have a total length of $N=13,421$ samples and a sampling period of 
%$\Delta t= 5,000 \text{yrs}$. The records and their corresponding RPs are shown in 
%Figure~\ref{fig_tauRR_cenogrid_RPs} together with remarkable climate transitions in the Cenozoic (detrended time series in Fig.~\ref{fig_time_series_cenogrid} in 
%Appendix~\ref{sec_appendix_cenogrid}). Among other things, the large-scale patterns in the RPs shown were used by \textcite{Westerhold2020} to classify major climate 
%states: Hothouse, Warmhouse, Coldhouse and Icehouse. 
%
%\begin{figure}[h!]
% \centering
% \includegraphics[width=\textwidth]{./figures/chapter06_tauRR/fig_tauRR_cenogrid_RPs}
% \caption{\textbf{A} $\delta^{13}$C and \textbf{B} $\delta^{18}$O isotope records and their corresponding RPs (\textbf{C+D}), obtained from the non-embedded time 
% series under a fixed recurrence threshold corresponding to 8\% global RR. These plots were used by \textcite{Westerhold2020} to classify the major climate states 
% Hothouse, Warmhouse, Cold- and Icehouse.
% }
%\label{fig_tauRR_cenogrid_RPs}
%\end{figure}
%
%
%While in Chapter \ref{chapter:MCDTS}, 
%Section \ref{sec_mcdts_cenogrid_prediction} predictions were made for the detrended $\delta^{13}$C record, here the focus is on the analysis of the characteristic time scales 
%in both, the detrended $\delta^{18}$O and the detrended $\delta^{13}$C isotope records. Specifically, we have computed spike powerspectra of the $\tau$-RR windowed over each record. 
%Figures~\ref{fig_tauRR_O18_pecuzal},~\ref{fig_tauRR_C13_pecuzal} show that major features of their linear counterparts, namely ordinary FFT spectrograms (A panels), are also present in 
%these spectrograms. But there is additional information in the spike spectrograms (B panels). Firstly, overall there seem to be less frequencies excited than in the FFT spectrograms, but these 
%excited frequencies are therefore more pronounced. This is especially apparent in the oxygen isotope record spectrogram (Fig.~\ref{fig_tauRR_O18_pecuzal}), in which the obliquity-related time 
%scale (\mytilde 41 kyrs) is very prominent from 17-0 Mio yrs BP in the spike spectrogram (panel B). Our spike spectrograms also support the general conlusion given by \textcite{Westerhold2020} 
%that ``in the Hothouse and Warmhouse climate states, as well as the first Coolhouse phase, eccentricity-related cycles (\mytilde 100 \& \mytilde 405 kyrs) dominate the CENOGRID records, while 
%after the increasing influence of highlatitude cooling and ice growth during the second Coolhouse phase, the obliquity-band response steadily increases after 13.9 Mio yrs BP (middle Miocene 
%climate transition) before dominating climate dynamics by \mytilde 5 Mio yrs BP (late Miocene–early Pliocene)''. The spike spectrograms indicate a strong obliquity influence already from 17 
%Mio yrs BP (the onset of the Miocene Climate optimum). Moreover, in the carbon isotope record based spike spectrogram (Fig.~\ref{fig_tauRR_C13_pecuzal}B) there is a strong influence of the 
%100-kyr-eccentricity cycle in the first Warmhouse phase from \mytilde 61-56 Mio yrs BP, which is not represented in the FFT-based spectrogram (panel A). 
%
%
%\begin{figure}
% \centering
% \includegraphics[width=\textwidth]{./figures/fig_tauRR_O18_pecuzal}
% \caption{Evolutionary spectrograms based on \textbf{A} FFT powerspectra of the detrended $\delta^{18}$O time series and \textbf{B} 
% on the spike powerspectra of the 
% $\tau$-RR of the PECUZAL embedded $\delta^{18}$O record. The spectrograms were computed with a $w = 1,000 (\hat{=}~\text{5-Myr})$ window 
% and a step $ws=1$. We used the 
% embedding computed in Chapter~\ref{chapter:MCDTS} (Section~\ref{sec_mcdts_cenogrid_prediction}, see Table~\ref{tab:cenogrid_embedding}) 
% and a recurrence threshold fixed to 
% 5\% global RR, in order to ensure comparability within the different windows \cite{kraemer2018} (Chapter~\ref{chapter:threshold}). 
% Dashed red horizontal lines indicate major climate transitions according to \textcite{Westerhold2020}.
% }
%\label{fig_tauRR_O18_pecuzal}
%\end{figure}
%
%\begin{figure}
% \centering
% \includegraphics[width=\textwidth]{./figures/fig_tauRR_C13_pecuzal}
% \caption{Same as in Figure \ref{fig_tauRR_O18_pecuzal}, but for the detrended $\delta^{13}$C isotope record.
% }
%\label{fig_tauRR_C13_pecuzal}
%\end{figure}
%
%The shown spike spectrograms in Figs.~\ref{fig_tauRR_O18_pecuzal},~\ref{fig_tauRR_C13_pecuzal} are based on the $\tau$-RR of the respective records, which have been time delay embedded by using 
%the PECUZAL algorithm \cite{Kraemer2021} (Chapter \ref{chapter:unified}). By construction, we would expect different spike spectrograms, when different embeddings are used. This is indeed the case 
%as shown in Figures~\ref{fig_tauRR_O18_mcdts},~\ref{fig_tauRR_C13_mcdts},~\ref{fig_tauRR_O18_ne},~\ref{fig_tauRR_C13_ne} (Appendix \ref{appendix_tauRR_spectrograms}), which also show spike 
%spectrograms, but for an embedding using MCDTS-R-MSE (Chapter~\ref{chapter:MCDTS}, see Table~\ref{tab:cenogrid_embedding}) and for the non-embedded time series. Even though differences are 
%discernible, the mentioned overall conclusion given by \textcite{Westerhold2020} also holds for these cases.

\section{Conclusion}\label{sec_tau_rr_conclusion}

A novel type of powerspectrum, the \textit{inter spike spectrum}, has been proposed. The method decomposes any arbitrary signal into basis functions which consist of (lagged) Dirac 
combs (DC) of different inter-spike period. The loading for each period is obtained by a regularized regression, which promotes sparsity in its solution (we chose LASSO). Since there are 
$M = \sum_{i=1}^{N}i$ basis functions for a signal of length $N$ the regression can get computationally intensive for $N>1,000$. When plotting the computed loadings as a function 
of the frequency (or period) the inter spike spectrum is obtained. This novel powerspectrum is superior to an ordinary FFT-based powerspectrum, when the signal has a spike-train-like 
appearance. The invention of the proposed method has been motivated by the idea of transforming $\tau$-recurrence rate signals ($\tau$-RR's) into their frequency domain. 
This general idea \cite{Zbilut2008} allows for a frequency analysis of high dimensional systems, because the RP is a representation of the system's state space trajectory.   
The $\tau$-RR of a recurrence plot (RP) usually has a spiky shape, especially for map-like data, and the inter spike spectrum can reliably reveal the system's dominant frequencies, 
which is not possible when Fourier transforming the $\tau$-RR or the underlying signal itself. Since the position of the peaks 
in the $\tau$-RR are not sensitive to noise, the corresponding inter spike spectrum also yields robust results in the presence of noise. We have successfully used the idea 
of transforming the $\tau$-RR for the detection of bifurcations in the Logistic map. By constructing appropriate surrogates of the inter spike spectra, and thus a null model, 
the number of significant peaks in the inter spike spectrum correlated well with the positive Lyapunov exponent. This measure was also able to resolve period-doubling bifurcations.

We could think of a broad range of applications of the proposed idea. The inter spike spectrum itself can serve as a valuable tool for the analysis of any sort of 
spike-train-like data. On the other hand, the inter spike spectrum of the $\tau$-RR of a signal can serve as a generalized, nonlinear frequency analysis tool for complex systems. 
When there is only a subset of state variables available, the state space has to be reconstructed as a pre-processing step. Recent findings \cite{Kraemer2021} show that 
this reconstruction process can be reliably automated and applied to multivariate data as well. This would allow for a ``running window'' approach, in order to detect transitions. 
Due to the mentioned computational constraints of our proposed method, a window size $w\leq 1,000$ would possibly suffice for most data, especially when it is map-like, i.e., not 
highly sampled.

\begin{acknowledgements}
This work has been financially supported by the German Research Foundation (DFG projects MA4759/8 and MA4759/9). All computations have been carried out in the Julia language 
and made use of the packages \textit{DynamicalSystems.jl} \cite{datseris2018} and \textit{DifferentialEquations.jl} \cite{rackauckas2017}.
\end{acknowledgements}

% Authors must disclose all relationships or interests that 
% could have direct or potential influence or impart bias on 
% the work: 
%
 \section*{Conflict of interest}

 The authors declare that they have no conflict of interest.

\clearpage
\appendix
\section{Inter spike spectra for noisy R\"ossler system}

\begin{figure}[h!]
 \centering
 \includegraphics[width=0.9\textwidth]{./figures/fig_tau_rr_example_roessler_noise}
 \caption{Same as in Fig.~\ref{fig_tau_rr_example_roessler}, but here with 5\% additive Gaussian white noise. The appearance of the inter spike spectra 
 in \textbf{J, K, L} and the Fourier spectra in \textbf{M, N, P} are unaffected by the additive noise noise.
 }
\label{fig_tau_rr_example_roessler_noise}
\end{figure}

\section{Exemplary models}

\subsection{Lorenz system}\label{sec_models_lorenz63}

The classical Lorenz-63 system \cite{lorenz1963} is defined as

\begin{equation}
\begin{array}{rcl}
\dot{x}&=&\sigma(y-x) \\
\dot{y}&=&x(r-z)-y \\
\dot{z}&=&xy - \beta z.
\end{array}
\label{eq_model_Lorenz63}
\end{equation}

\noindent For producing Fig.~\ref{fig_tau_rr_spectrum_example} we set the initial condition to $u_0=[0.0, 10.0, 0.0]$, used a sampling time of $\Delta t=0.01$ and discarded the first 
2,000 points of the integration as transients. The parameters have been set to 
$\sigma=10, \beta=8/3, \rho=28$ and we used a time series consisting of 6,000 samples.

\subsection{R\"ossler system}\label{sec_models_roessler}

The R\"ossler system \cite{roessler1976} is defined as
\begin{align}
\begin{array}{rcl}
\dot{x}&=&-y-z \\
\dot{y}&=&x+ay \\
\dot{z}&=&b+ z(x-c) .
\end{array}
\label{eq_model_roessler}
\end{align}
For producing Figs.~\ref{fig_tau_rr_example_roessler_noise},~\ref{fig_tau_rr_example_roessler} the initial condition for producing panels \textbf{A} \& \textbf{B} was set to $u_0=[.7, -1, 0.4]$ with a sampling time 
of $dt=0.05$ and in case of panel \textbf{C} $u_0=[-0.1242, -2.5415, 0.2772]$ with a sampling time of $dt=0.1$. The first $5,000$ samples were discarded as transients and trajectories of length $N=5,000$ were 
obtained from which we computed the RPs and the corresponding $\tau$-RR's. For the inter spike spectra only the first $1,000$ values of the $\tau$-RR's were considered.

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
%\bibliography{}   % name your BibTeX data base

% Non-BibTeX users please use
\bibliographystyle{spbasic} 
\bibliography{inter_spike_bib}

\end{document}
% end of file