### 2022-11-16 Manuscript/entropy-1919488  proof/entropy-1919488.tex
proof

%  LaTeX support: latex@mdpi.com 
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
\documentclass[entropy,article,accept,pdftex,moreauthors]{Definitions/mdpi} 
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%=================================================================
% MDPI internal commands
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2022}
\copyrightyear{2022}
\externaleditor{Academic Editor: Firstname Lastname}
\datereceived{31 August 2022} 
%\daterevised{} % Only for the journal Acoustics
\dateaccepted{14 November 2022} 
\datepublished{} 
%\datecorrected{} % Corrected papers include a "Corrected: XXX" date in the original paper.
%\dateretracted{} % Corrected papers include a "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption

% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
\newcommand\hmmax{0}
\newcommand\bmmax{0}

\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\setlength{\marginparwidth}{2cm}
\usepackage{changes}
\usepackage{bbm}
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines
%\usepackage{mathptmx}
\usepackage{amssymb}
\usepackage{statmath}
\usepackage{enumitem}
\usepackage{mathtools}
\usepackage{siunitx}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
%
% please place your own definitions here and don't use \def but
\newcommand{\mytilde}{\raise.17ex\hbox{$\scriptstyle\mathtt{\sim}$}}

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Spike Spectra for Recurrences}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Spike Spectra for Recurrences}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0002-9943-5391} % Add \orcidA{} behind the author's name
\newcommand{\orcidauthorB}{0000-0001-6643-9508} % Add \orcidB{} behind the author's name
\newcommand{\orcidauthorC}{0000-0003-1437-7039} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{\hl{K. Hauke Kraemer} %MDPI: Please carefully check the accuracy of names and affiliations.
 $^{1,}$*\orcidA{}, Frank Hellmann $^{1}$, Mehrnaz Anvari $^{1}$\orcidB{}, \hl{J\"urgen Kurths} %MDPI: The name of this author is different from the submitting system. Please confirm which one is correct.
 $^{1,2,\highlighting{3} %MDPI: 
}$ and Norbert Marwan $^{1,2,\highlighting{4} %MDPI: 
,}$*\orcidC{}}

%\longauthorlist{yes}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{K. Hauke Kraemer, Frank Hellmann, Mehrnaz Anvari, J\"urgen Kurths and Norbert Marwan}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Kraemer, K.H.; Hellmann, F.; Anvari, M.; Kurths, J.; Marwan, N.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Potsdam Institute for Climate Impact Research, Member of the Leibniz Association, 14473 Potsdam, Germany; \hl{hellmann@pik-potsdam.de (F.H.); anvari@pik-potsdam.de (M.A.); kurths@pik-potsdam.de (J.K.)} %MDPI: We added the email addresses here according to the submitting system. Please confirm.
\\
$^{2}$ \quad Institute of Physics and Astronomy, University of Potsdam, 14476 Potsdam, Germany\\
\hl{$^{3}$} %MDPI: The affiliation numbers should appear in numerical order. Please check the suggested changes.
 \quad Institute of Physics, Humboldt Universit\"at zu Berlin, 12489 Berlin, Germany\\
\hl{$^{4}$} \quad Institute of Geosciences, University of Potsdam, 14476 Potsdam, Germany}

% Contact information of the corresponding author
\corres{Correspondence: hauke\_kraemer@hotmail.com (K.H.K.); marwan@pik-potsdam.de (N.M.)}

% Current address and/or shared authorship
%\firstnote{Current address: Affiliation 3.} 
%\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

% Abstract (Do not insert blank lines, i.e. \\)

\abstract{In recurrence analysis, the $\tau$-recurrence rate encodes the periods of the cycles of the underlying high-dimensional time series. 
It, thus, plays a similar role to the autocorrelation for scalar time-series in encoding temporal correlations. 
However, its Fourier decomposition does not have a clean interpretation. 
Thus, there is no satisfactory analogue to the power spectrum in recurrence analysis. 
We introduce a novel method to decompose the $\tau$-recurrence rate using an over-complete basis of Dirac combs together with sparsity regularization. 
We show that this decomposition, the \textit{inter-spike spectrum}, naturally provides an analogue to the power spectrum for recurrence analysis in the sense that it reveals the dominant periodicities of the underlying time series. 
We show that the inter-spike spectrum correctly identifies patterns and transitions in the underlying system in a wide variety of examples and is robust to measurement noise.}

% Keywords
\keyword{decomposition; frequency analysis; recurrence analysis; bifurcations}
\PACS{05.45.Tp; 05.90.+m; 89.90.+n; 02.70.Uu; 05.10.Ln; 05.45.-a; 05.45.Ac}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\section{Introduction}\label{sec_tau_rr_intro}

The dynamics of complex systems as provided by measured time series usually show complicated and chaotic patterns.
Quantifying their recurrence features is a powerful way to describe them and to infer information about 
the type of dynamics, stability, regime changes, or couplings and synchronisation \cite{marwan2007,marwan2008epjst,webber2015}.
Even more challenging are signals which do have a heavy tailed-distribution or appear as a spike-train,
e.g., neuron firings \cite{Dummer2014,Orcioni2020}, heart beat variability \cite{marwan2002herz}, 
or extreme flood events \cite{banerjee2021}.
Deriving useful information from spike-train signals or inter-spike time series is an important
topic in data analysis in many scientific fields \cite{Kajikawa2005,Dummer2014,Orcioni2020,Canale2021}.

Recurrence plots (RPs) provide a vivid representation of complex dynamics $\vec{x}_i$ stemming from potentially high dimensional systems \cite{marwan2007}
\begin{equation}\label{eq_rp_definition}
R_{i,j}(\varepsilon) = \Theta\left(\varepsilon - D_{i,j}\right) 
= \Theta\left(\varepsilon - \| \vec{x}_i - \vec{x}_j\|\right), \qquad \vec{x} \in \mathbb{R}^d, \quad i,j \in [1,\ldots, N],
\end{equation}
where $\mathbf{R}$ \hl{is} %MDPI: There is no bold in other places in the text, please confirm whether the bold here needs to be retained
 the recurrence matrix, $\vec{x}_i$ is the state vector at time 
$t = \Delta t \cdot i$ ($\Delta t$ the sampling time), $N$ is the number of
sampling points (or length of data series), and $d$ is the dimension of the system.
The crucial free parameter $\varepsilon$ is the recurrence threshold, determining what is a recurrence
and, thus, the visible structures in the RP. It can be chosen such that the recurrence rate 
$RR(\varepsilon)=N^{-2}\sum_{i,j}^N R_{i,j}(\varepsilon)$ exhibits a certain value \cite{kraemer2018}.
The simple idea to track recurring states of the $d$-dimensional trajectory $\vec{x}_i$ of the system under study not only allows for a beneficial visualization of the dynamics, 
but also for its 
quantification, using certain structures in the RP, such as diagonal or vertical lines \cite{marwan2007}.

Some of these recurrence quantification measures, the entropy of diagonal lines and the entropy of 
recurrence times, can be related to the basic characteristics of complex systems, such as Kolmogorov--Sinai entropy \cite{march2005,baptista2010}. However, these quantifiers have a free parameter, the minimal considered line length, and 
are usually biased, due to the finite size of the RP and thickened diagonal lines, which need to be corrected \cite{Kraemer2019}. Moreover, the mentioned statistics cannot account for 
changing regular (non-chaotic) dynamics, such as period-doubling bifurcations.

A rather simple idea is to look at the $\tau$-recurrence rate of the RP ($\tau$-RR, Equation~(\ref{eq_tau_rr}))~\cite{marwan2002pla,Zbilut2008}.
This is the density of recurrence points along the diagonals of the recurrence matrix, as a function of the distance $\tau$ (sampling units) to the main diagonal:
\begin{equation}\label{eq_tau_rr}
\tau\text{-RR}(\varepsilon) = RR(\tau, \varepsilon) = \frac{1}{N-\tau} \sum_{i=1}^{N-\tau	} R_{i,i+\tau}.
\end{equation}
$\tau$-RR serves as an estimator for the probability that the system recurs after time $\tau \Delta t$, with $\Delta t$ being the sampling time of the trajectory 
$\vec{x}_i = \vec{x}(\Delta t \cdot i),\, i=1,\ldots,N$. 
It represents the period length of cycles in the data (Figure~\ref{fig_tau_rr_spectrum_example}D).

\citet{Zbilut2008} found that $\tau$-RR could be interpreted as analogous to the auto-correlation function $C(\tau)$ and, hence, via the Wiener--Khinchin theorem, provide an analogue,
``generalized'', spectral density. This is reasonable, since the average distances for a given lag $\tau$ 
\begin{equation}
\overline{D}(\tau) = \frac{1}{N-\tau}\sum_{i=1}^{N-\tau} D_{i, i+\tau}
\end{equation}
can be directly read from the distance matrix $\mathbf{D}$ and are also preserved in its thresholded version $\tau$-RR. There are clear advantages to recurrence-derived spectral density, 
i.e., Fourier transforming (FT) $\tau$-RR (Figure~\ref{fig_tau_rr_spectrum_example}D,E), instead 
of $C(\tau)$. There are no assumptions for stationarity or sampling when constructing a RP.
Furthermore, since an RP can represent high-dimensional dynamics and its $\tau$-RR serves as a plug-in for $C(\tau)$, the correlation structures of higher dimensional spaces can be read from 
the recurrence-derived Fourier~spectrum.\\
\vspace{-6pt}
\begin{figure}[H]
 %\centering
 \includegraphics[width=0.9\textwidth]{./figures/fig_tau_rr_spectrum_example}
 \caption{\hl{Schematic} %MDPI: Please change the hyphen (-) into minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1".
 illustration of a $\tau$-recurrence-rate-based spectrum. (\textbf{A}) $x$-component time series of the Lorenz63-System (Equation~\eqref{eq_model_Lorenz63}) and 
 (\textbf{B}) its corresponding Fourier power spectrum. 
 (\textbf{C})~Reconstructed state space portrait from the time series shown in (\textbf{A}) using PECUZAL time-delay \hl{embedding} %MDPI: Please ensure that permission has been obtained and there is no copyright issue. If copyright is needed, please provide a citation in the following format: "Reprinted/adapted with permission from Ref. [XX]. Copyright year, copyright owner's name". More details on "Copyright and Licensing" are available via the following link: https://www.mdpi.com/ethics#10.
 \cite{Kraemer2021}. 
 (\textbf{D}) Subset of the recurrence plot and the corresponding $\tau$-recurrence rate obtained from the state space trajectory in (\textbf{C}). 
(\textbf{E}) Fourier Power spectrum obtained from the $\tau$-recurrence rate (subset shown in panel (\textbf{D}\hl{))} %MDPI: Please ensure that permission has been obtained and there is no copyright issue. If copyright is needed, please provide a citation in the following format: "Reprinted/adapted with permission from Ref. [XX]. Copyright year, copyright owner's name". More details on "Copyright and Licensing" are available via the following link: https://www.mdpi.com/ethics#10.
 \cite{Zbilut2008}. (\textbf{D},\textbf{E}) show the results of a part of the time series, 
 which is highlighted in pink in (\textbf{A}).
 }\label{fig_tau_rr_spectrum_example}
\end{figure}

However, the interpretation of this generalized power spectral density is unclear, and it is typically hard to interpret. $\tau$-RR directly encodes the periodicity of the underlying signal; in contrast, it is unclear what, if any, interpretation the ``power'' contained in a particular frequency mode of the $\tau$-RR should be. Whenever $\tau$-RR is a spike-train-like signal, which it is in most cases (see Figure~\ref{fig_tau_rr_spectrum_example}) especially for 
map-data (low-resolution data), an FT of such a signal leads to a spike-train-like image in the frequency domain (e.g.,~\cite{Schild1982,Cordoba1989}, see 
Figure~\ref{fig_tau_rr_spectrum_example}E). Thus, how to extract meaningful information about the systems' state space trajectory is not intuitive.

For example, consider the signal we would like to analyze 
(e.g., the $\tau$-RR of a system) to be a Dirac comb (DC) with inter-spike period $T_\text{is}$: 
\begin{equation}
\text{DC}_{\text{is}}(t) = \sum_{k=-\infty}^{\infty} \delta(t-kT_\text{is}),
\label{eq_dirac_comb}
\end{equation}
i.e., a series of Dirac delta functions for a period $T_\text{is}$. There is only one single period, $T_\text{is}$, in this signal (Figure~\ref{fig_tau_rr_dirac_comb}A,D); 
in principle, we would strive for a single peak in the frequency domain of this signal at frequency $f=1/T_\text{is}$. The Fourier spectrum does not meet this expectation 
and instead of a single frequency, there are many excited frequencies (Figure~\ref{fig_tau_rr_dirac_comb}B,E). This is because the Fourier components constructively contribute
to every frequency $1/T_\text{is}$; therefore, $\text{DC}_{\text{is}}(t)$ coincides with its own Fourier transform up to a factor $1/T_\text{is}$ \cite{Norden1998}. ({\hl{This can also } %MDPI: footnote is not allowed in this journal, we moved here, please confirm
be observed for neuron spike trains, e.g., \cite{Orcioni2020,Biagetti2017}.})

\begin{figure}[H]
% \centering
 \includegraphics[width=0.9\textwidth]{./figures/fig_tau_rr_dirac_comb}
 \caption{The transformation of a Dirac comb (series of Dirac delta functions) with a single inter-spike period $T_{\text{is}}=100$ ($\widehat{=}f=0.01$) 
 into the frequency domain. 
 (\textbf{A}) Dirac Comb (DC) with equal amplitudes and
 (\textbf{B}) its FFT-based power spectral density.
 (\textbf{C}) Proposed inter-spike spectrum of the signal in (\textbf{A})~showing a single frequency, which corresponds to the inter-spike period $T_{\text{is}}$ ($f=0.01$).
 (\textbf{D}) DC with randomly chosen amplitudes and same $T_{\text{is}}$ as in (\textbf{A}), and
 (\textbf{E}) is its FFT-based power spectral density. 
 (\textbf{F}) Proposed inter-spike spectrum of the signal in (\textbf{D}) showing a single frequency, which corresponds to the expected inter-spike period $T_{\text{is}}$ ($f=0.01$). 
 Inter-spike spectra were obtained with a LASSO regression and a regularization threshold corresponding to $\rho=0.9$ accordance of the signals in (\textbf{A},\textbf{D}) and 
 its re-composed signals (c.f. Section~\ref{sec_tau_rr_method}). 
}
\label{fig_tau_rr_dirac_comb}
\end{figure}

In this article, we propose a new way of decomposing a spike-train-like signal into periodic components as 
an alternative to the RP-based method suggested in \cite{Zbilut2008} or Fourier-based spike-train power 
spectra \cite{Dummer2014}. This novel \textit{inter-spike spectrum} does not show resonance 
behavior of the signal's inherent inter-spike frequencies in such a way that the harmonics of these frequencies are also excited (Figure~\ref{fig_tau_rr_dirac_comb}C,F). Section \ref{sec_tau_rr_method} explains the idea. Note that this approach can be used to decompose arbitrary signals, and is not specific to $\tau$-RR. However, the more spiky the signal is, the more useful our new approach is compared to 
FT. In Section \ref{sec_tau_rr_application}, we demonstrate the usage of inter-spike spectrum when transforming the $\tau$-RR of a system under study. In this case, the inter-spike spectrum 
can unravel characteristic time scales of high dimensional systems, which is not possible when using an FT. Finally, in Section \ref{sec_tau_rr_conclusion} our results are summarized.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Method}\label{sec_tau_rr_method}

To obtain the inter-spike spectrum, the signal, in our case the $\tau$-RR, is decomposed into a set of appropriate basis functions. 
The general idea common to many methods is that the sum of these weighted functions can approximate a finite signal to a sufficient degree. The weights (in some contexts also called modes or loadings) corresponding to the individual basis functions must be determined.
A number of decomposition techniques based on different sets of basis functions exist, e.g., 
trigonometric functions (Fourier and wavelet analysis \cite{Bracewell1986}), 
eigenvectors of the corresponding covariance matrix (principal component analysis \cite{Hotelling1933} and related techniques) 
or intrinsic mode functions (empirical mode decomposition and Hilbert spectrum \cite{Norden1998}).
These methods typically share the property that the basis functions form a complete basis, and
the set of basis functions is linearly independent; thus, the weights are uniquely determined.

Here, we propose the use of Dirac combs (DC) with 
different inter-spike periods as basis functions, as shown in Equation~\eqref{eq_dirac_comb}. Let $s(t_i)$ be the min-max-normalized signal we want to transform in terms of length $N$ and 
$t_i=i\cdot \Delta t,~i=1,\ldots,N$, where $\Delta t$ denotes the sampling time and $s(t_i) \in [0,\ 1]\ \forall\ i$. In the following sections, we label this time series as a $(1\times N)$-dimensional 
vector $\vec{s}$. 
First, $\tilde{N}$ different DCs of length $N$ are constructed with inter-spike periods $T_\text{is} \in [1,\ldots,\tilde{N}]$ and $\tilde{N}=\ceil*{N/2}+1$. Second, in order to account 
for possible phase shifts of 
these basis functions occurring in $\vec{s}$, each of these $\tilde{N}$ different DCs also need to be shifted one step further $T_\text{is}-1$ times. This leaves us with a total number of 
$M = \sum_{i=1}^{\tilde{N}}i$ 
basis functions which we can arrange as rows of a $(M\times N)$-sized matrix $\bfX$
(Figure~\ref{fig_tau_rr_basis_functions} illustrates the described procedure) as follows:
\begin{align}
\label{eq_basis_matrix} 
\bfX_{i,j} &= \sum_{k=0}^{N}\delta\bigl(j-1-kT_\text{is}(i)-i+T_\text{is}(i) \bigr), \quad i=1,\dots,\ceil*{N/2}+1, \quad j=1,\dots,N\\
\label{eq_T_is} T_\text{is}(i)&= n, \quad \text{such that}\quad n:~ \frac{n(n-1)}{2} + 1 \leq i <  \frac{n(n+1)}{2} + 1,\quad n \in \mathbb{N}_+.
\end{align}

Note that due to the shifting of each of the basis functions of the inter-spike period $T_\text{is}$, $\bfX$ is no longer linear-independent. 
Furthermore, there will be identical basis functions, which do not allow for an unambiguous inter-spike period, 
if we would include all $N$ possible inter-spike periods for a signal of length $N$ instead of $\ceil*{N/2}+1$ (Figure~\ref{fig_tau_rr_basis_functions}A).
The reason is that in contrast to a trigonometric decomposition, where the Nyquist frequency marks a lower bound for the corresponding wave period, 
the maximum considered inter-spike period is bounded by $T_{\text{is}}^{\text{max}} = \ceil*{N/2}+1$ (schematically illustrated in 
Figure~\ref{fig_tau_rr_basis_functions}B).

Eventually, an under-determined linear system
\begin{equation}
\bfX^T \boldsymbol{\beta}=\vec{s}
\label{eq_linear_system}
\end{equation}  
has to be solved for $\boldsymbol\beta$, with the $(M\times 1)$-sized vector carrying the loadings we are interested in. Of the variety of algorithms which can solve this problem, we are 
particularly interested in those solutions which promote sparsity in $\boldsymbol\beta$, since our goal is to decompose the signal $\vec{s}$ into a minimal number of basis 
functions (for an excellent overview of the topic we refer readers to \citet{Brunton2019}). In this paper, we either use the \textit{least absolute shrinkage and selection operator} 
(LASSO) \cite{Tibshirani1996} or a \textit{sequentially threshold least squares} (STLS) regression~\cite{Brunton2016,Brunton2019} to obtain a solution $\hat{\boldsymbol\beta}$. Any other 
sparse regression technique can be used. Finally, we group loadings 
which correspond to basis functions with the same period 
$T_\text{is}$ into $\hat{\boldsymbol\beta}_f$ and obtain the 
\textit{inter-spike spectrum} by simply plotting $\hat{\boldsymbol\beta}_f$ as a function of the frequency $f=T_\text{is}^{-1}$, with $T_\text{is}=\Delta t, 2\Delta t, \dots, (\ceil*{N/2}+1) \Delta t$ 
(Figure~\ref{fig_tau_rr_dirac_comb}C, F).

\begin{figure}[H]
%\centering
\includegraphics[width=0.9\textwidth]{./figures/fig_tau_rr_basis_functions}
\caption{(\textbf{A}) Example of a full set of basis functions for an input signal of length $N=4$, aligned in the matrix $\bfX'$. Inter-spike periods $T_{\text{is}}$ larger 
than $\ceil*{N/2}+1$ lead to redundant basis functions (i.e., repeated lines in $\bfX$, red sheared) or basis functions, which cannot be uniquely assigned to a certain inter-spike period 
(blue sheared). $T_{\text{is}}$ for each row can be obtained by Equation~\eqref{eq_T_is}. (\textbf{B}) The final set of unambiguous, but still linearly dependent, basis functions aligned in the matrix $\bfX$.} \label{fig_tau_rr_basis_functions}
\end{figure}

When introducing a sparsity regularization, we forsake the ability to fully reconstruct the signal from the decomposition in favor of identifying the most prominent spikes, and thus, the most important periodicities, more accurately. This choice is parametrized by a regression regularization parameter $\alpha$. We only obtain a solution to Equation~(\ref{eq_linear_system}) with $\alpha = 0$ as well as a full reconstruction of the signal. Our general approach to fixing $\alpha$ is to preserve most of the original signal in the sense of matching a given (Pearson) correlation coefficient $\rho$ between the original signal $\vec{s}$ and the reconstruction. The chosen regression method (in our case, LASSO or STLS) will determine how sensitive the obtained inter-spike spectrum is with respect to the regularization of the 
selected $\rho$ (Figure~\ref{fig_regularization}). Regarding this, in most of our applications LASSO yields more robust results, while STLS is usually faster in terms of computation. 
We will discuss this in more detail in Section \ref{sec_tau_rr_discussion}.
The proposed decomposition does not show a leakage effect by construction. However, when introducing sparsity regularization we forsake the ability to fully reconstruct the signal from the decomposition in favor of identifying the most important periodicities more clearly. This choice is mediated by a regression regularization parameter $\alpha$. Only for $\alpha = 0$ do we obtain a solution to Equation~(\ref{eq_linear_system}) and thus a full reconstruction of the signal. Our general approach to fixing $\alpha$ is to preserve most of the original signal in the sense of matching a given (Pearson) correlation coefficient $\rho$ between the original signal $\vec{s}$ and the reconstruction. The chosen regression method (in our case, LASSO or STLS) will determine how sensitive the obtained inter-spike spectrum is with respect to the regularization achieved by a 
selected $\rho$ (Figure~\ref{fig_regularization}). Regarding this, in most of our applications LASSO yields more robust results, while STLS is usually faster in terms of computation. %Attention AE: Please check repetition in previous pragraphs%
We will discuss this in more detail in Section \ref{sec_tau_rr_discussion}.

%
%In addition to the computational complexity of solving Equation~\eqref{eq_linear_system}, the solving algorithm must also deal with the free choice of the regression regularization parameter $\alpha$, 
%which determines the sparsity of the regression coefficient vector $\hat{\boldsymbol\beta}$. Obviously, there will be numerous peaks in the inter-spike spectrum of a given signal, when $\alpha$ is small, 
%since $\alpha=0$ corresponds to the least squares solution and the corresponding $\hat{\boldsymbol\beta}$ is an exact solution of Equation~(\ref{eq_linear_system}) and not sparse. On the other hand, 
%only a sparse number of non-zero loadings in $\hat{\boldsymbol\beta}$, i.e., peaks in the inter-spike spectrum, are obtained when $\alpha$ is sufficiently large. 
%The actual implementation of the regularization depends on the regression method chosen (in our case, LASSO or STLS), and consequently a given value of $\alpha$ will generally result in 
%different sparse loading vectors $\hat{\boldsymbol\beta}$. Therefore, in this paper we select $\alpha$ such that the re-composed signal $\tilde{\vec{s}}=\bfX^T\hat{\boldsymbol\beta}$ matches a 
%given (Pearson) correlation coefficient $\rho$ between the original signal $\vec{s}$ and itself.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}\label{sec_tau_rr_application}

We demonstrate the use of the inter-spike spectrum in combination with $\tau$-RR, as outlined in 
Section \ref{sec_tau_rr_intro}, on several interesting research questions. The procedure takes place as follows:
\begin{itemize}[noitemsep]
\item[(1)] Compute an RP of the trajectory of the system, Equation~\eqref{eq_rp_definition}. If only univariate data are available, perform a state space reconstruction for obtaining the trajectory first.
\item[(2)] Compute the $\tau$-RR of the RP, as shown in Equation~\eqref{eq_tau_rr}.
\item[(3)] Transform the $\tau$-RR into the proposed inter-spike spectrum, see Section \ref{sec_tau_rr_method}.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Period Estimation for Different Dynamics in the R \"ossler System}
First, we consider the R\"ossler system (Equation~\eqref{eq_model_roessler}, shown in Appendix \ref{sec_models_roessler}), 
in three different dynamical setups. We use the proposed inter-spike spectrum to
identify the type of dynamics.
We set the parameters as $b=2$, $c=4$ and analyze period-2 limit cycle dynamics ($a=0.36$ in Figure~\ref{fig_tau_rr_example_roessler}A,D,G,J), 
period-3 limit cycle dynamics ($a=0.41$ in Figure~\ref{fig_tau_rr_example_roessler}B,E,H,K) and chaotic dynamics ($a=0.428$ in Figure~\ref{fig_tau_rr_example_roessler}C,F,I,L).

\begin{figure}[H]
% \centering
 \includegraphics[width=0.76\textwidth]{./figures/fig_tau_rr_example_roessler}
 \caption{\hl{Inter-spike spectra} %MDPI: Please use a consistent numbering style for the subfigures throughout the paper. P is after N;  Please change the hyphen (-) into minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1".
 of the $\tau$-RR of the R\"ossler system in three different dynamical regimes with parameters $b=2$, $c=4$. 
 (\textbf{A}) Trajectory of the system in a period-2 (parameter $a=0.36$), (\textbf{B})~in a period-3 (parameter $a=0.41$) and 
 (\textbf{C}) in a chaotic regime (parameter $a=0.428$). 
\mbox{ (\textbf{D}--\textbf{F})}~The corresponding RPs, obtained by using a recurrence threshold corresponding to a 10\% global 
 recurrence rate for (\textbf{D},\textbf{E}) and 5\% for (\textbf{F}). 
  (\textbf{G}--\textbf{I}) $\tau$-RRs of the shown RPs. 
  (\textbf{J}--\textbf{L}) The proposed inter-spike spectra of the $\tau$-RRs shown in panels (\textbf{G}--\textbf{I}). Spectra were obtained with a LASSO regression and a regularization threshold 
  corresponding to $\rho=0.95$ accordance of $\tau$-RRs and re-composed signals. The distance ratio of the peaks reflect the limit cycle dynamic.  
  (\textbf{M},\textbf{N},\textbf{P}) Fourier power spectra of the $\tau$-RRs shown in panels (\textbf{G}--\textbf{I}). }
\label{fig_tau_rr_example_roessler}
\end{figure}

The inter-spike spectra unravel the specific dynamics, which are also apparent in the state space portraits (Figure~\ref{fig_tau_rr_example_roessler}A--C) and in the 
$\tau$-RRs (Figure~\ref{fig_tau_rr_example_roessler}G--I). The proposed idea is also 
robust to noise (see Figure~\ref{fig_tau_rr_example_roessler_noise} in the Appendix~\ref{appc}). This is because the peaks of $\tau$-RR are insensitive to noise. 
Figure \ref{fig_roessler_peaks} illustrates that 
while the peak shape does change in the presence of noise, its position does not, and this is what the inter-spike spectrum encrypts 
(see Figure~\ref{fig_roessler_spectra_distances} in the Appendix~\ref{appc} for further analysis).

\begin{figure}[H]
 %\centering
 \includegraphics[width=0.9\textwidth]{./figures/fig_roessler_different_noise_levels_peaks}
 \caption{Peak positions of the obtained inter-spike spectra of the $\tau$-RRs for additive noise levels up to 50\% for the discussed R\"ossler dynamics, 
 (\textbf{A}) period-2 limit-cycle, (\textbf{B}) period-3 limit-cycle and (\textbf{C})~chaos. The size of the plotted markers scale with the detected peak height. The noise-free spectra are shown in Figure~\ref{fig_tau_rr_example_roessler}J--L and 
 an example of these spectra with 5\% additive noise is shown in Figure~\ref{fig_tau_rr_example_roessler_noise}J--L. Spectra were obtained with a LASSO regression and a regularization threshold 
  corresponding to $\rho=0.95$ accordance of $\tau$-RRs and re-composed signals.}
\label{fig_roessler_peaks}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Bifurcations in the Logistic Map}

We consider the Logistic map $x_{n+1}=r\cdot x_n \left( 1-x_n \right)$ for changing the control parameter $r$. We vary $r$ from $r=3.4$ to $r=4$ in steps of $0.001$. For 
each setting of $r$ 
\begin{itemize}[noitemsep]
\item[(1)] A time series of length $N=201$ is computed with a random initial condition \mbox{$u_0 \in [0,\ 1]$}, neglecting the first $1,000$ samples as transients;
\item[(2)] A \textls[-15]{total of $100$ iterative Amplitude Adjusted Fourier Transform (iAAFT) surrogates}~\cite{Schreiber1996,Schreiber2000} are computed;
\item[(3)] The time series and its iAAFT surrogates are embedded in a 2-dimensional state space using a time delay of unity;
\item[(4)] The two-dimensional trajectories RPs, as shown in Equation~\eqref{eq_rp_definition}, are computed under a threshold $\varepsilon=0.05$,
\item[(5)] $\tau$-RR, as shown in Equation~\eqref{eq_tau_rr}, is computed from the RP of the signal and from the RPs of the surrogates;
\item[(6)] Inter-spike spectra are obtained from $\tau$-RR of the signal and from the $\tau$-RRs of the surrogates, as can be seen in Section \ref{sec_tau_rr_method};
\item[(7)] Finally, from the distribution of the surrogate inter-spike spectra, the 95th percentile is computed. The peaks of the inter-spike spectrum of the signal which exceed 
this percentile are counted. 
\end{itemize}

\hl{In this example} %MDPI: we add indent here, please confirm
, the null hypothesis for constructing the surrogate data is that the data stem from a process which yields the same auto-correlation, and
hence, the same Fourier power spectral density, and the same 
amplitude distribution. We consider the number of significant peaks in the inter-spike spectrum with
respect to the control parameter in order to distinguish the corresponding dynamics (Figure~\ref{fig_tau_rr_logistic}C).
A correlation with the positive Lyapunov exponent (Figure~\ref{fig_tau_rr_logistic}A) is discernible 
($\rho_{\text{Pearson}}(\text{Lyapunov})=0.72$). This analysis can handle period-doubling since it measures the dominant cycles via the inter-spike spectrum. 
However, whenever the periods of the new cycles coincide with integer multiples of the periods of already existing cycles, the approach cannot detect 
period doubling. A similar case can be observed in Figure~\ref{fig_tau_rr_example_roessler}J,K, where the number of peaks does not change, but  the mutual distance does.

A less computationally intensive approach is to compute surrogates for $\tau$-RR analytically, rather than computing an RP and its $\tau$-RR for each iAAFT surrogate of the 
time series. This translates into a null hypothesis 
that the $\tau$-RR and its corresponding inter-spike spectrum stem from an RP of a random signal. In this case, the probability of finding a black point in the RP can be obtained 
from a binomial distribution with the probability parameter $p$ set to the recurrence rate of the RP of the signal. In this way, $100$ surrogate $\tau$-RRs are computed in step (5). 
The results are even slightly better compared to the ones obtained from the iAAFT surrogates (Figure~\ref{fig_tau_rr_logistic}B, $\rho_{\text{Pearson}}(\text{Lyapunov})=0.81$). 
The first instance of period doubling at $r \approx 3.458$ cannot be detected by any of the surrogates.

The described procedure does work well for map data, because most often the $\tau$-RR for those kind of data reveals a ``spiky enough'' nature. 
On the contrary, highly sampled (flow-) data often yield not as 
spiky $\tau$-RRs; therefore, the number of significant peaks in the inter-spike spectrum may not be sensitive enough to detect period-doubling bifurcations. Moreover the sensitivity of the 
inter-spike spectrum in detecting regime shifts also depends on the critical regularization threshold. Nevertheless, the 
according inter-spike spectra still reveal important information (Figure~\ref{fig_tau_rr_example_roessler}) and practitioners can design appropriate quantifying statistics based 
on these spectra which suit the research task.

\begin{figure}[H]
 %\centering
 \includegraphics[width=\textwidth]{./figures/fig_tau_rr_logistic}
 \caption{(\textbf{A}) Bifurcation diagram and Lyapunov exponent and of the Logistic map as a function of the control parameter $r$.
 (\textbf{B}) Number of significant peaks ($\alpha=0.05$) in the inter-spike spectrum of the $\tau$-RR and its Pearson correlation coefficient to the Lyapunov exponent shown in (\textbf{A})
 (white noise surrogates). 
 (\textbf{C}) Same as (\textbf{B}), but for iterative Amplitude Adjusted Fourier Transform (iAAFT) \hl{surrogates} %MDPI: Please ensure that permission has been obtained and there is no copyright issue. If copyright is needed, please provide a citation in the following format: "Reprinted/adapted with permission from Ref. [XX]. Copyright year, copyright owner's name". More details on "Copyright and Licensing" are available via the following link: https://www.mdpi.com/ethics#10.
 \cite{Schreiber1996,Schreiber2000}. To obtain the inter-spike spectra, we used 
 a LASSO regression and a regularization threshold corresponding to the $\rho=0.95$ accordance of $\tau$-RRs and re-composed signals.  
}
\label{fig_tau_rr_logistic}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Inter-Spike Spectra of Power Grid Frequency Data}\label{sec_power_grid}

Power grids are large, synchronized, complex networks whose stable functioning is indispensable for modern societies.
To maintain the stability of a power grid, the balance between energy consumption and energy generation must be ensured. 
In an AC-power grid, the grid frequency is an observable variable that reflects how well this balance is satisfied. 
In this process, the grid frequency and its deviations from the nominal frequency are continuously recorded and monitored by 
the grid operators (in Europe and many parts of the world this is $50$ \si{Hz} or $60$ \si{Hz} in America and, for example, southern Japan).
For example, if there is more (less) demand than supply, the network frequency decreases (increases) compared to the nominal frequency \cite{kundur1994power}.

The frequency variations can include other information, such as the functionality of control systems \cite{gorjao2020data}, the effect of fluctuations in
renewable energies (REs), demands on the grid \cite{anvari2020stochastic} and, moreover, the effect of regular dispatches due to the trading market 
\cite{meyer2020identifying}. The latter induce periodic frequency jumps. Here we look at the frequency time series for the Great Britain (GB) and Continental Europe 
(CE) (Appendix \ref{sec_power_grid_appendix} and Figure~\ref{fig_power_grid_time_series}A,C). Clear jumps every 30 and 60 min are discernible and quantitatively reflected in the corresponding 
autocorrelations (Figure~\ref{fig_power_grid_time_series}B,D). Furthermore, the autocorrelation of the CE frequency time series shows regular peaks every $15$ \si{min} 
(see Figure~\ref{fig_power_grid_time_series}B). These peaks are caused by a mismatch of power supply and demand \cite{weissbach2009high} during dispatches. 
In most electricity grids the operation of dispatchable power plants is scheduled in 1-h blocks, where additional (shorter) $30$ and $15$ \si{min} intervals
might exist.

\begin{figure}[H]
 \centering
 \includegraphics[width=\textwidth]{./figures/fig_power_grid_spectra}
 \caption{Averaged evolutionary Fourier power spectra of recorded power grid frequency time series of (\textbf{A}) Central Europe (CE) and (\textbf{B}) Great Britain (GB) (see 
 Figure~\ref{fig_power_grid_time_series} in Appendix~\ref{sec_power_grid_appendix}). 
 The corresponding averaged inter-spike spectra of the according $\tau$-RRs, Equation~\eqref{eq_tau_rr}, are shown in (\textbf{C},\textbf{D}), respectively. Vertical red dashed lines correspond to 
 $15$, $30$ and $60$ \si{min}. For technical details on the calculation of the spectra shown, i.e, the preprocessing and window size being used, the reader is referred to Appendix~\ref{sec_power_grid_appendix}.}  
\label{fig_power_grid_spectra}
\end{figure}

The evolutionary Fourier frequency spectrum for the Central European data in \linebreak \mbox{Figure~\ref{fig_power_grid_spectra}A}, however, does not display sharp peaks exactly at $15$, $30$ and $60$ \si{min}, which may partly 
be explained by the leakage effect (for technical details on the calculation of 
the spectra shown, the reader is referred to Appendix~\ref{sec_power_grid_appendix}). More clearly, the $30$ and $60$ \si{min} 
peaks are split into two adjacent peaks and the local minimum in between these ``double''-peaks correspond to the exact times. The Great Britain counterpart in Figure~\ref{fig_power_grid_spectra}B 
has sharp peaks at $15$ and $30$ \si{min} and also ``local predecessor peaks'' for $30$ and $60$ \si{min} at the same positions as in panel A. 
The fact that sharp peaks are seen here using the same window size and sampling frequency as in the case of the Central European data shows that the leakage effect alone is not the cause of 
the peak splitting, as shown in Figure~\ref{fig_power_grid_spectra}A.

In contrast, the $15$ \si{min} peak is completely absent from the inter-spike spectra of the $\tau$-RRs of the frequency data (Figure~\ref{fig_power_grid_spectra}C,D), but these show sharp peaks 
at $30$ and $60$ \si{min} (there is no leakage effect due to the proposed decomposition technique). 
In front of these peaks, smaller local peaks can be seen, which correspond to the local peaks of the Fourier power spectra at $28$ and $55$ \si{min}, respectively.
Moreover, there is an additional peak at 
$40$~\si{min} for both datasets, which is absent in the Fourier spectrum and which is not a multiple of the missing $15$ \si{min} peak. The position and magnitude of the peaks in the 
inter-spike spectra are robust to the chosen recurrence threshold, the regression method and its regularization as well as the sampling time of the original signal.

We interpret the results presented as follows. The missing $15$ \si{min} in the inter-spike spectra is due to the much stronger autocorrelation at $30$ and $60$ \si{min} 
(see Figure~\ref{fig_power_grid_time_series}B,D); because these periods are integer multiples of $15$ \si{min}, the inter-spike spectra are not able to detect them and the 
sparse regression ``drags'' the $15$ \si{min} periods into the mentioned $30$ and $60$ \si{min} peaks. Moreover, the inter-spike spectra, unlike the Fourier spectra, demonstrate sharp peaks exactly 
at $30$ and $60$ \si{min}, i.e., during dispatches (not valid in case of the GB dataset). Finally,  we found a clear sharp peak at $40$ \si{min} which can occur 
because of any regular controls in a power grid, and can indicate the need to develop the existing stochastic processes to model the power grid frequency more precisely \cite{gorjao2020data}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evolutionary Inter-Spike Spectra of Earth'S Orbit Data}

When applying the proposed inter-spike spectrum to the $\tau$-RR of a time series we expect additional frequency/period information, due to the fact that the recurrence plot (RP), 
Equation~\eqref{eq_rp_definition}, and its corresponding $\tau$-RR, as shown in Equation~\eqref{eq_tau_rr}, visualize the trajectory of the embedded time series in an embedding space of a higher dimension. 
However, given sufficient embedding of the time series, we would also expect that major frequencies/periods of the non-embedded time series are incorporated in the RP, its $\tau$-RR, 
and eventually in the inter-spike spectrum of the $\tau$-RR. In order to demonstrate this, we apply the inter-spike spectrum to the freely available eccentricity time series of \mbox{\citet{Laskar2011}}. 
This astronomical computation of the orbital motion of the Earth (here, we focus on the eccentricity only) has a clear expectation value of the incorporated frequencies/periods. The three leading 
eccentricity cycles of 405 kyr period, 95 kyr period and 124 kyr period are well known in palaeoclimate studies \cite{Laskar2004,Westerhold2020}. Our aim in this section is to show that the inter-spike spectrum 
of the $\tau$-RR of the embedded eccentricity time series will reflect these cycles in a similar fashion as the Fourier power spectral density of the non-embedded eccentricity time series. We will further 
show that Fourier-transforming the $\tau$-RR instead of applying the proposed inter-spike spectrum will lead to non-satisfying results, since the spiky $\tau$-RR excites a variety of harmonics in 
the corresponding Fourier spectrum (cf. Section \ref{sec_tau_rr_intro}). We use a time series which covers the past \mytilde $67$ Mio.~years (Myr), with a total length of $N=13,421$ samples and a 
sampling period of $\Delta t= 5000$.

First, we compute an evolutionary short time FT using a windowsize of $ws=1000$ samplepoints ($\equiv 5$ Myr) shifted by unity and a Hamming window. The spectrogram reveals the expected periods mentioned, which are highlighted and clearly visible (Figure~\ref{fig_laskar_spectra}A).

 \hl{Then}, we construct the inter-spike spectrogram of the $\tau$-RR by first determining an appropriate embedding. By using a recent tree-embedding ansatz \cite{Kraemer2022}, we minimize the 
\textit{false nearest neighbor} statistic \cite{hegger1999} and use the continuity statistic \cite{pecora2007} for potential delays. Eventually, we obtain $8$-dimensional embedding 
with embedding delays $\tau =[0, 18, 36, 48, 62, 101, 114]$ (in sampling units) for the entire time series. Similar to the preceding approach we embed the time series with these embedding parameters 
on windows of size $ws=1,200$ and a unity shift. ({\hl{We use a slightly} larger window here than in the short FT, because the embedding causes a ``loss'' of data points and we want to 
cover similar time spans.}) The RPs are computed on these embedded trajectories with a fixed recurrence threshold corresponding to a $10\%$ global recurrence rate \cite{kraemer2018} 
and the inter-spike spectra of the corresponding $\tau$-RRs is obtained by using STLS regression (see Section \ref{sec_tau_rr_method}) and a regularization threshold corresponding to 
$\rho=0.9$ accordance of $\tau$-RRs and re-composed signals. We only use the first $200$ data points of the $\tau$-RR (covering a time span of $1$ Myr). 
The spectrogram also highlights the $95$, $124$ and $405$ kyr periods as expected (Figure~\ref{fig_laskar_spectra}B). 
Additional power is distributed in the harmonics at $190$ and $248$ kyr periods. Finally, the standard FT with a Hamming window of the same $\tau$-RRs used to obtain the inter-spike 
spectrogram in panel B yields a smeared spectrogram which does not reflect the expected periods, but rather suffers from the spike train behavior, i.e., many excited harmonics, of the 
FT described in Section \ref{sec_tau_rr_intro} and Figure~\ref{fig_tau_rr_dirac_comb}. The shown results are robust to a change of the embedding parameters and windowsizes. However, a low 
regularization threshold smears the clear spectrogram in panel B.

\begin{figure}[H]
 %\centering
 \includegraphics[width=\textwidth]{figures/fig_laskar_spectra}
 \caption{(\textbf{A}) Evolutionary Fourier power spectra of eccentricity time series. (\textbf{B}) Inter-spike spectrogram of the $\tau$-recurrence rate of the eccentricity time 
 series and (\textbf{C}) its Fourier spectrogram. Horizontal black dashed lines highlight the analytically expected orbital periods of $405$, $124$ and $95$ kyrs. 
 For comparability, in all cases the spectra aligned in the columns of the shown plots are normalized to probabilities (sum of unity for 
 each power spectrum). For further computational details, please refer to the main text.}  
\label{fig_laskar_spectra}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}\label{sec_tau_rr_discussion}

We successfully used the idea of transforming the $\tau$-RR for the detection of bifurcations in the Logistic map. By constructing appropriate surrogates of the inter-spike spectra, 
and, thus, a null model, the number of significant peaks in the inter-spike spectrum correlated well with the positive Lyapunov exponent. This measure was also able to resolve 
period-doubling bifurcations. However, we have to admit that by using this method, the detection of a bifurcation is only possible when the additional period(s) is not an integer multiple of the former period(s). 
This behavior is described in the application to the R\"ossler system, where we explicitly showed the different inter-spike spectra for period-2, period-3, and chaotic dynamics.
Further development might potentially incorporate the mutual distance of peaks in the spectrum for a better correlation to the Lyapunov exponent. The inter-spike spectra of power grid 
	frequency data illustrate that our proposed method may serve as a valuable source of information in addition to a standard Fourier analysis. Last but not least, we showed that our approach incorporates frequencies, which are apparent in the Fourier spectrum of the signal, by applying the method to analytically derived eccentricity data, where the dominant frequencies are well 
known.

We discuss some more technical details in the following sections, which will also affect any application of the proposed method. 
First of all, the number of required basis functions 
$M = \sum_{i=1}^{\ceil*{N/2}+1}i$ for an input signal of length $N$ is the crucial bottleneck of this approach, which is why it does not show good scaling behavior. The subsequent sparse 
regression, therefore, becomes computationally intensive for $N>1000$. Depending on the memory of the computer used, input signals $N>2000$ usually do not work. This means that signals 
often need to be downsampled as a preprocessing step (e.g., see Appendix \ref{sec_power_grid_appendix}). Second, the regularization parameter $\alpha$ for the regression is a crucial free parameter. 
As described in Section~\ref{sec_tau_rr_method}, our idea in this paper is to select the $\alpha$ such that the re-composed signal $\tilde{\vec{s}}=\bfX^T\hat{\boldsymbol\beta}$ matches a 
given (Pearson) correlation coefficient $\rho_{\vec{s},\tilde{\vec{s}}}$ between the original signal $\vec{s}$ and itself. This ensures that $\alpha$ adjusts itself to the data as well as to the 
used regression method. We found that this increases the comparability of different spectra, especially when performing a running window approach in order to obtain an evolutionary spectrogram 
(Figure~\ref{fig_laskar_spectra}). 
However, the two different sparse regression algorithms we encountered in this article (LASSO and STLS) yield different results for the same desired $\rho_{\vec{s},\tilde{\vec{s}}}$. 
Even if the spectra obtained in this way look qualitatively similar, they are not always quantitatively similar. The reason for this 
is that $\rho_{\vec{s},\tilde{\vec{s}}}$ is not a smooth function of $\alpha$ in case of STLS, due to the hard-thresholding involved \cite{Brunton2016}, which is shown in Figure~\ref{fig_regularization}.
Third, when adopting our idea of applying the inter-spike spectrum to the $\tau$-recurrence rate of the signals state space trajectory, the embedding process induces additional free parameters.  
This is not a drawback of the proposed decomposition method, but rather a drawback of applying this technique to the $\tau$-recurrence rate of the system, which was the main motivation for 
developing the proposed method. As a very last remark, we draw attention to the fact that sparse regression can be transformed into sparse logistic regression when the signal we would like to transform is binary.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}\label{sec_tau_rr_conclusion}

A novel decomposition technique is proposed that yields the \textit{inter-spike spectrum}. The method decomposes any arbitrary signal into basis functions which consist of (lagged) Dirac 
combs (DC) of a different inter-spike period. The loading for each period is obtained by a regularized regression, which promotes sparsity in its solution. We chose LASSO or a sequentially 
thresholded least squares regression STLS in this letter. Since there are 
$M = \sum_{i=1}^{\ceil*{N/2}+1}i$ basis functions for a signal of length $N$, the regression can become computationally intensive for $N>1000$. When plotting the computed loadings as a function 
of the period (or frequency), the inter-spike spectrum is obtained. A disadvantage is that the transformation is not invertible. An advantage is that there is no leakage effect.
Although this novel spectrum is superior to an ordinary FFT-based power spectrum when the signal has a spike-train-like appearance, the authors suggest that this method should 
be considered 
as an additional source of information but not as a substitute for ordinary Fourier analysis. 
Due to the sparse regression underlying the method, there is no unique inverse of the transformation and the regularization parameter plays a crucial role and determines the 
appearance of the obtained inter-spike spectrum. Moreover, similar to the Nyquist frequency barrier in the Fourier Transform which sets a lower bound for the corresponding wave period, here 
the maximum considered inter-spike period is bounded by $T_{\text{is}}^{\text{max}} = \ceil*{N/2}+1$.

The creation of the proposed method was by the idea of transforming $\tau$-recurrence rate signals ($\tau$-RRs) into their frequency domain. 
This general idea \cite{Zbilut2008} facilitates a frequency analysis of high dimensional systems, because the RP is a representation of the system's state space trajectory.   
The $\tau$-RR of a recurrence plot (RP) usually has a spiky shape, especially for map-like data, and the inter-spike spectrum can reliably reveal the system's dominant frequencies, 
which is not possible when Fourier transforming the $\tau$-RR or the underlying signal itself. Since the position of the peaks 
in the $\tau$-RR are not sensitive to noise, the corresponding inter-spike spectrum also yields robust results in the presence of~noise.

There is a broad range of applications for the proposed idea. The inter-spike spectrum itself can serve as a valuable tool for the analysis of any sort of 
spike-train-like data. On the other hand, the inter-spike spectrum of the $\tau$-RR of a signal can serve as a generalized, nonlinear frequency analysis tool for complex systems. 
When there is only a subset of state variables available, the state space has to be reconstructed as a pre-processing step. Recent findings \cite{Kraemer2021,Kraemer2022} show that 
this reconstruction process can be reliably automated and applied to multivariate data as well. This would allow for a ``running window'' approach, in order to detect transitions. 
Due to the mentioned computational constraints of our proposed method, a window size $w\leq 1000$ would possibly suffice for most data, especially when it is map-like, i.e., not 
highly sampled.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{Conceptualization, K.H.K. and F.H.; methodology, K.H.K., F.H. and N.M.; software, K.H.K. and F.H.; validation, K.H.K., F.H., M.A. and N.M.; formal analysis, K.H.K.; data curation, K.H.K. and M.A.; writing---original draft preparation, K.H.K.; writing---review and editing, K.H.K., F.H., M.A. and N.M.; visualization, K.H.K.; supervision, N.M.; project administration, J.K. All authors have read and agreed to the published version of the manuscript.}

\funding{This work has been financially supported by the German Research Foundation (DFG projects MA4759/8 and MA4759/9).}

\institutionalreview {\hl{ }}%{In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study. You might choose to exclude this statement if the study did not require ethical approval. Please note that the Editorial Office might ask you for further information. Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans. OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals. OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.}

\informedconsent {\hl{ }}%{Any research article describing a study involving humans should contain this statement. Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans. You might also choose to exclude this statement if the study did not involve humans. Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves). Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.}

\dataavailability{The study that we present here is available as a fully reproducible code base
``repository will be published and cited here, when accepted'' and the method
will be available in the Julia language, ``package will be published and cited here, 
when accepted'' and as a MATLAB\textsuperscript{\textregistered} toolbox 
``Toolbox will be published and cited here, when accepted''.}

\acknowledgments{
All computations have been carried out in MATLAB\textsuperscript{\textregistered} and the Julia language 
and we made use of the packages \textit{DynamicalSystems.jl} \cite{datseris2018}, \textit{DifferentialEquations.jl} \cite{rackauckas2017}, \textit{Distances.jl} and 
\textit{OptimalTransport.jl} \cite{Zhang2022}.
}

\conflictsofinterest{The authors declare that they have no conflict of interest.} 
\vspace{-6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
\appendixtitles{yes} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixstart
\appendix

\section{Exemplary Models}

\subsection{Lorenz System}\label{sec_models_lorenz63}

\hl{The} classical Lorenz-63 system \cite{lorenz1963} is defined as

\begin{equation}
\begin{array}{rcl}
\dot{x}&=&\sigma(y-x) \\
\dot{y}&=&x(r-z)-y \\
\dot{z}&=&xy - \beta z.
\end{array}
\label{eq_model_Lorenz63}
\end{equation}
{For} producing Figure~\ref{fig_tau_rr_spectrum_example}, we set the initial condition to $u_0=[0.0, 10.0, 0.0]$, used a sampling time of $\Delta t=0.01$ and discarded the first 
2000 points of the integration as transients. The parameters have been set to 
$\sigma=10, \beta=8/3, \rho=28$ and we used a time series consisting of 6000 samples.

\subsection{R \"ossler System}\label{sec_models_roessler}

\hl{The} R\"ossler system \cite{roessler1976} is defined as
\begin{align}
\begin{array}{rcl}
\dot{x}&=&-y-z \\
\dot{y}&=&x+ay \\
\dot{z}&=&b+ z(x-c) .
\end{array}
\label{eq_model_roessler}
\end{align}
{For} producing Figures~\ref{fig_tau_rr_example_roessler_noise} and ~\ref{fig_tau_rr_example_roessler}, the initial condition for producing panels \textbf{A} \& \textbf{B} was set to $u_0=[0.7, -1, 0.4]$ with a sampling time 
of $dt=0.05$ and in case of panel \textbf{C}, $u_0=[-0.1242, -2.5415, 0.2772]$ with a sampling time of $dt=0.1$. The first $5000$ samples were discarded as transients and trajectories of length $N=5000$ were 
obtained from which we computed the RPs and the corresponding $\tau$-RRs. For the inter-spike spectra, only the first $1000$ values of the $\tau$-RRs were considered.

\section{Power Grid Frequency Time Series}\label{sec_power_grid_appendix}

\textls[-15]{The raw frequency time series have a length of $\tilde{N}_{\text{CE}}=58,752,000$ and} \mbox{$\tilde{N}_{\text{GB}}=31,622,400$} with sampling times $\Delta t_{\text{CE}}=0.2$ \si{s} and 
$\Delta t_{\text{GB}}=1$ \si{s}, respectively. GB frequency data were measured during 2016, and were obtained from \cite{GB}. CE frequency data were from 2017 and were obtained from \cite{haehne2018footprint}.

We downsampled these time series to a sampling time of $\Delta t_{\text{CE}}= \Delta t_{\text{GB}}=20$ \si{s}, which led to total time series lengths of 
$N_{\text{CE}}=587,520$ and $N_{\text{GB}}=1,581,120$ which we used for further analysis.

\begin{figure}[H]
% \centering
 \includegraphics[width=\textwidth]{./figures/fig_power_grid_time_series}
 \caption{Subset of the frequency time series of (\textbf{A}) Central Europe \hl{(CE)} %MDPI: Please ensure that permission has been obtained and there is no copyright issue. If copyright is needed, please provide a citation in the following format: "Reprinted/adapted with permission from Ref. [XX]. Copyright year, copyright owner's name". More details on "Copyright and Licensing" are available via the following link: https://www.mdpi.com/ethics#10.
 \cite{haehne2018footprint} and (\textbf{C}) Great Britain \hl{(GB)} %MDPI: 
 \cite{GB} along with their autocorrelation functions in (\textbf{B},\textbf{D}), respectively. The normalized autocorrelations in (\textbf{B},\textbf{D}) were computed with the entire time series, whereas the subsets shown in (\textbf{A},\textbf{C}) 
 contain only $1441$ samples.}
\label{fig_power_grid_time_series}
\end{figure}

In this analysis, we divided the time series into 
non-overlapping blocks of length $N_{\text{block}}=4320$, covering a time span of $24$ h. For each time series block we computed Fourier spectra and recurrence plots (RPs), 
Equation~\eqref{eq_rp_definition}, along with their $\tau$-RRs, Equation~(\ref{eq_tau_rr}). The RPs were obtained from a uniform 5-dimensional time-delay embedding of each time series with 
the delay set to the first minimum of the mutual information \cite{fraser1986,hegger1999} and a fixed recurrence threshold corresponding to an 8\% recurrence rate was used in order 
to ensure comparability \cite{kraemer2018}. The first $600$ data points of the $\tau$-RRs were used to obtain inter-spike spectra with a LASSO regression and a 
regularization threshold corresponding to $\rho=0.95$ accordance of $\tau$-RRs and re-composed signals. The spectra shown in all panels of Figure~\ref{fig_power_grid_spectra} 
are the averages over all blocks, following Meyer et al~\cite{meyer2020identifying}.

\begin{figure}[H]
% \centering
 \includegraphics[width=\textwidth]{./figures/fig_power_grid_fft_full}
 \caption{Fourier spectra of the frequency entire time series from (\textbf{A}) Central Europe (CE), \mbox{$N_{\text{CE}}=587,520$} and (\textbf{B}) Great Britain (GB), $N_{\text{GB}}=1,581,120$. A subset of the underlying time series is shown in Figure~\ref{fig_power_grid_time_series}.}
\label{fig_power_grid_fft_full}
\end{figure}

\section{Inter-Spike Spectra for Noisy R \"ossler System}\label{appc}

\begin{figure}[H]
 %\centering
 \includegraphics[width=0.93\textwidth]{./figures/A3A}

 \caption{\textit{Cont}.}
\label{fig:petit3leaks3D}
\end{figure}

\begin{figure}[H]\ContinuedFloat

 \includegraphics[width=0.93\textwidth]{./figures/A3B}
 \caption{\hl{Same as in} %MDPI: please add explanation for A--I
 Figure~\ref{fig_tau_rr_example_roessler}, but here with 5\% additive Gaussian white noise on each component $x$, $y$ and $z$. The appearance of the inter-spike spectra 
 in (\textbf{J}--\textbf{L}), and the Fourier spectra in (\textbf{M},\textbf{N},\textbf{P}) are unaffected by the additive noise.}
\label{fig_tau_rr_example_roessler_noise}
\end{figure}

\vspace{-6pt}

\begin{figure}[H]
 %\centering
 \includegraphics[width=0.9\textwidth]{./figures/fig_roessler_spectra_distances}
 \caption{Wasserstein distances for the obtained inter-spike spectra of the $\tau$-RRs for additive noise levels of up to 50\% for the discussed R\"ossler dynamics, 
 (\textbf{A}) period-2 limit-cycle, (\textbf{B}) period-3 limit-cycle and (\textbf{C}) chaos. For each noise level the obtained spectrum is compared to the noise-free spectrum (these are shown 
 in Figure~\ref{fig_tau_rr_example_roessler}J--L for a regularization threshold corresponding to $\rho=0.95$) by computing its Wasserstein distance. Inter-spike spectra were obtained with a 
 LASSO regression and three different regularization thresholds corresponding to $\rho=0.9$ (blue), $\rho=0.95$ (orange) and $\rho=0.99$ (green) accordance of $\tau$-RRs and re-composed signals. 
 Up to a noise level of 20\% the distances are small and rather constant. The fluctuations depend on the chosen regularization threshold as well as on the underlying dynamics.}
\label{fig_roessler_spectra_distances}
\end{figure}

\section{Regularizationparameters for Different Sparse Regression Techniques}\label{sec_regularization_appendix}

\begin{figure}[H]
 \centering
 \includegraphics[width=\textwidth]{./figures/fig_regularization}
 \caption{Pearson correlation coefficient $\rho_{\vec{s},\tilde{\vec{s}}}$ between the input signal signal $\vec{s}$ and the re-composed signal $\tilde{\vec{s}}=\bfX^T\hat{\boldsymbol\beta}$ as a 
 function of the regularization parameter $\alpha$ for (\textbf{A}) LASSO and (\textbf{B}) sequentially thresholded least squares (STLS) regression (see also Section \ref{sec_tau_rr_method}). 
 The input time series is of length $N=200$ and stems from the $\tau$-RR of R\"ossler system in regular dynamics.}
\label{fig_regularization}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{adjustwidth}{-\extralength}{0cm}
%\printendnotes[custom] % Un-comment to print a list of endnotes

\reftitle{References}

\begin{thebibliography}{999}

\bibitem[Marwan {et~al.}(2007)Marwan, Romano, Thiel, and Kurths]{marwan2007}
Marwan, N.; Romano, M.C.; Thiel, M.; Kurths, J.
\newblock {Recurrence Plots for the Analysis of Complex Systems}.
\newblock {\em Phys. Rep.} {\bf 2007}, {\em 438},~237--329.
\newblock {https://doi.org/10.1016/j.physrep.2006.11.001}.

\bibitem[Marwan(2008)]{marwan2008epjst}
Marwan, N.
\newblock {A Historical Review of Recurrence Plots}.
\newblock {\em Eur. Phys. J. Spec. Top.} {\bf 2008}, {\em
  164},~3--12.
\newblock
  {https://doi.org/10.1140/epjst/e2008-00829-1}.

\bibitem[{Webber, Jr.} and Marwan(2015)]{webber2015}
{Webber, Jr.}, C.L.; Marwan, N.
\newblock {\em {Recurrence Quantification Analysis---Theory and Best
  Practices}}; Springer: Cham, Switzerland, 2015; p.~421.
\newblock
  {https://doi.org/10.1007/978-3-319-07155-8}.

\bibitem[Dummer {et~al.}(2014)Dummer, Wieland, and Lindner]{Dummer2014}
Dummer, B.; Wieland, S.; Lindner, B.
\newblock {Self-consistent determination of the spike-train power spectrum in a
  neural network with sparse connectivity}.
\newblock {\em Front. Comput. Neurosci.} {\bf 2014}, {\em
  8},~104.
\newblock
 {https://doi.org/10.3389/fncom.2014.00104}.

\bibitem[Orcioni {et~al.}(2020)Orcioni, Paffi, Apollonio, and
  Liberti]{Orcioni2020}
Orcioni, S.; Paffi, A.; Apollonio, F.; Liberti, M.
\newblock {Revealing Spectrum Features of Stochastic Neuron Spike Trains}.
\newblock {\em Mathematics} {\bf 2020}, {\em 8},~1011.
\newblock
  {https://doi.org/10.3390/math8061011}.

\bibitem[Marwan {et~al.}(2002)Marwan, Wessel, Meyerfeldt, Schirdewan, and
  Kurths]{marwan2002herz}
Marwan, N.; Wessel, N.; Meyerfeldt, U.; Schirdewan, A.; Kurths, J.
\newblock {Recurrence Plot Based Measures of Complexity and its Application to
  Heart Rate Variability Data}.
\newblock {\em {Phys. Rev. E}} {\bf 2002}, {\em 66},~026702.
\newblock
 {https://doi.org/10.1103/PhysRevE.66.026702}.

\bibitem[Banerjee {et~al.}(2021)Banerjee, Goswami, Hirata, Eroglu, Merz,
  Kurths, and Marwan]{banerjee2021}
Banerjee, A.; Goswami, B.; Hirata, Y.; Eroglu, D.; Merz, B.; Kurths, J.;
  Marwan, N.
\newblock Recurrence analysis of extreme event-like data.
\newblock {\em Nonlinear Process. Geophys.} {\bf 2021}, {\em
  28},~213--229.
\newblock
  {https://doi.org/10.5194/npg-28-213-2021}.

\bibitem[Kajikawa and Hackett(2005)]{Kajikawa2005}
Kajikawa, Y.; Hackett, T.A.
\newblock {Entropy analysis of neuronal spike train synchrony}.
\newblock {\em J. Neurosci. Methods} {\bf 2005}, {\em 149},~90--93.
\newblock
  {https://doi.org/10.1016/j.jneumeth.2005.05.011}.

\bibitem[Canale {et~al.}(2021)Canale, Lijoi, Nipoti, and
  Pr{\"{u}}nster]{Canale2021}
Canale, A.; Lijoi, A.; Nipoti, B.; Pr{\"{u}}nster, I.
\newblock {Inner spike and slab Bayesian nonparametric models}.
\newblock {\em Econom. Stat.} {\bf 2021}.
\newblock
  {https://doi.org/10.1016/j.ecosta.2021.10.017}.

\bibitem[Kraemer {et~al.}(2018)Kraemer, Donner, Heitzig, and
  Marwan]{kraemer2018}
Kraemer, K.H.; Donner, R.V.; Heitzig, J.; Marwan, N.
\newblock Recurrence threshold selection for obtaining robust recurrence
  characteristics in different embedding dimensions.
\newblock {\em Chaos: Interdiscip. J. Nonlinear Sci.} {\bf
  2018}, {\em 28},~085720.
\newblock
  {https://doi.org/10.1063/1.5024914}.

\bibitem[March {et~al.}(2005)March, Chapman, and Dendy]{march2005}
March, T.K.; Chapman, S.C.; Dendy, R.O.
\newblock Recurrence plot statistics and the effect of embedding.
\newblock {\em Phys. D} {\bf 2005}, {\em 200},~171--184.
\newblock
  {https://doi.org/10.1016/j.physd.2004.11.002}.

\bibitem[Baptista {et~al.}(2010)Baptista, Ngamga, Pinto, Brito, and
  Kurths]{baptista2010}
Baptista, M.S.; Ngamga, E.J.; Pinto, P.R.F.; Brito, M.; Kurths, J.
\newblock {Kolmogorov-Sinai entropy from recurrence times}.
\newblock {\em Phys. Lett. A} {\bf 2010}, {\em 374},~1135--1140.
\newblock
 {https://doi.org/10.1016/j.physleta.2009.12.057}.

\bibitem[Kraemer and Marwan(2019)]{Kraemer2019}
Kraemer, K.H.; Marwan, N.
\newblock Border effect corrections for diagonal line based recurrence
  quantification analysis measures.
\newblock {\em Phys. Lett. A} {\bf 2019}, {\em 383},~125977.
\newblock
 {https://doi.org/10.1016/j.physleta.2019.125977}.

\bibitem[Marwan and Kurths(2002)]{marwan2002pla}
Marwan, N.; Kurths, J.
\newblock {Nonlinear analysis of bivariate data with cross recurrence plots}.
\newblock {\em {Phys. Lett. A}} {\bf 2002}, {\em 302},~299--307.
\newblock
  {https://doi.org/10.1016/S0375-9601(02)01170-2}.

\bibitem[Zbilut and Marwan(2008)]{Zbilut2008}
Zbilut, J.P.; Marwan, N.
\newblock The Wiener–Khinchin theorem and recurrence quantification.
\newblock {\em Phys. Lett. A} {\bf 2008}, {\em 372},~6622--6626.
\newblock
 {https://doi.org/10.1016/j.physleta.2008.09.027}.

\bibitem[Kraemer {et~al.}(2021)Kraemer, Datseris, Kurths, Kiss,
  Ocampo-Espindola, and Marwan]{Kraemer2021}
Kraemer, K.H.; Datseris, G.; Kurths, J.; Kiss, I.Z.; Ocampo-Espindola, J.L.;
  Marwan, N.
\newblock A unified and automated approach to attractor reconstruction.
\newblock {\em New J. Phys.} {\bf 2021}, {\em 23},~033017.
\newblock
  {https://doi.org/10.1088/1367-2630/abe336}.

\bibitem[Schild(1982)]{Schild1982}
Schild, D.
\newblock An Efficient Method for the Fourier Transform of a Neuronal Spike
  Train.
\newblock {\em Int. J. Neurosci.} {\bf 1982}, {\em
  17},~179--182,
\newblock
  {https://doi.org/10.3109/00207458208985921}.

\bibitem[C{\'o}rdoba(1989)]{Cordoba1989}
C{\'o}rdoba, A.
\newblock Dirac combs.
\newblock {\em Lett. Math. Phys.} {\bf 1989}, {\em 17},~191--196.
\newblock
  {https://doi.org/10.1007/BF00401584}.

\bibitem[Huang {et~al.}(1998)Huang, Shen, Long, Wu, Shih, Zheng, Yen, Tung,
  and Liu]{Norden1998}
Huang, N.E.; Shen, Z.; Long, S.R.; Wu, M.C.; Shih, H.H.; Zheng, Q.; Yen, N.C.;
  Tung, C.C.; Liu, H.H.
\newblock The empirical mode decomposition and the Hilbert spectrum for
  nonlinear and non-stationary time series analysis.
\newblock {\em Proc. R. Soc. London. Ser. Math. Phys. Eng. Sci.} {\bf 1998}, {\em
  454},~903--995.
\newblock
 {https://doi.org/10.1098/rspa.1998.0193}.

\bibitem[Biagetti {et~al.}(2017)Biagetti, Crippa, Orcioni, and
  Turchetti]{Biagetti2017}
Biagetti, G.; Crippa, P.; Orcioni, S.; Turchetti, C.
\newblock Homomorphic Deconvolution for MUAP Estimation From Surface EMG
  Signals.
\newblock {\em IEEE J. Biomed. Health Inform.} {\bf 2017},
  {\em 21},~328--338.
\newblock
  {https://doi.org/10.1109/JBHI.2016.2530943}.

\bibitem[Bracewell and Bracewell(1986)]{Bracewell1986}
Bracewell, R.N.; Bracewell, R.N.
\newblock {\em The Fourier Transform and its Applications};
  McGraw-Hill: New York, NY, USA, 1986; Volume 31999.

\bibitem[Hotelling(1933)]{Hotelling1933}
Hotelling, H.
\newblock Analysis of a complex of statistical variables into principal
  components.
\newblock {\em J. Educ. Psychol.} {\bf 1933}, {\em 24},~417.

\bibitem[Brunton and Kutz(2019)]{Brunton2019}
Brunton, S.L.; Kutz, J.N.
\newblock {\em Data-Driven Science and Engineering: Machine Learning, Dynamical
  Systems, and Control}; Cambridge University Press: Cambridge, UK, 2019.
\newblock
  {https://doi.org/10.1017/9781108380690}.

\bibitem[Tibshirani(1996)]{Tibshirani1996}
Tibshirani, R.
\newblock Regression Shrinkage and Selection Via the Lasso.
\newblock {\em J. R. Stat. Soc. Ser. B} {\bf 1996}, {\em 58},~267--288.
\newblock
 {https://doi.org/10.1111/\linebreak j.2517-6161.1996.tb02080.x}.

\bibitem[Brunton {et~al.}(2016)Brunton, Proctor, and Kutz]{Brunton2016}
Brunton, S.L.; Proctor, J.L.; Kutz, J.N.
\newblock Discovering governing equations from data by sparse identification of
  nonlinear dynamical systems.
\newblock {\em Proc. Natl. Acad. Sci. USA} {\bf 2016},
  {\em 113},~3932--3937.
\newblock
  {https://doi.org/10.1073/pnas.1517384113}.

\bibitem[Schreiber and Schmitz(1996)]{Schreiber1996}
Schreiber, T.; Schmitz, A.
\newblock Improved Surrogate Data for Nonlinearity Tests.
\newblock {\em Phys. Rev. Lett.} {\bf 1996}, {\em 77},~635--638.
\newblock
 {https://doi.org/10.1103/PhysRevLett.77.635}.

\bibitem[Schreiber and Schmitz(2000)]{Schreiber2000}
Schreiber, T.; Schmitz, A.
\newblock Surrogate time series.
\newblock {\em Phys. D: Nonlinear Phenom.} {\bf 2000}, {\em 142},~346--382.
\newblock
 {https://doi.org/10.1016/S0167-2789(00)00043-9}.

\bibitem[Kundur {et~al.}(1994)Kundur, Balu, and Lauby]{kundur1994power}
Kundur, P.; Balu, N.J.; Lauby, M.G.
\newblock {\em Power System Stability and Control}; McGraw-hill: New
  York, NY, USA, 1994; Volume~7.

\bibitem[Gorj{\~a}o {et~al.}(2020)Gorj{\~a}o, Anvari, Kantz, Beck, Witthaut,
  Timme, and Sch{\"a}fer]{gorjao2020data}
Gorj{\~a}o, L.R.; Anvari, M.; Kantz, H.; Beck, C.; Witthaut, D.; Timme, M.;
  Sch{\"a}fer, B.
\newblock Data-driven model of the power-grid frequency dynamics.
\newblock {\em IEEE Access} {\bf 2020}, {\em 8},~43082--43097.

\bibitem[Anvari {et~al.}(2020)Anvari, Gorj{\~a}o, Timme, Witthaut,
  Sch{\"a}fer, and Kantz]{anvari2020stochastic}
Anvari, M.; Gorj{\~a}o, L.R.; Timme, M.; Witthaut, D.; Sch{\"a}fer, B.; Kantz,
  H.
\newblock Stochastic properties of the frequency dynamics in real and synthetic
  power grids.
\newblock {\em Phys. Rev. Res.} {\bf 2020}, {\em 2},~013339.

\bibitem[Meyer {et~al.}(2020)Meyer, Anvari, and Kantz]{meyer2020identifying}
Meyer, P.G.; Anvari, M.; Kantz, H.
\newblock Identifying characteristic time scales in power grid frequency
  fluctuations with DFA.
\newblock {\em Chaos: Interdiscip. J. Nonlinear Sci.} {\bf
  2020}, {\em 30},~013130.

\bibitem[Wei{\ss}bach and Welfonder(2009)]{weissbach2009high}
Wei{\ss}bach, T.; Welfonder, E.
\newblock High frequency deviations within the European power system: Origins
  and proposals for improvement.
\newblock In Proceedings of the 2009 IEEE/PES Power Systems Conference and
  Exposition, Seattle, WA, USA, 15--18 March 2009; pp.~1--6.

\bibitem[{Laskar, J.} {et~al.}(2011){Laskar, J.}, {Fienga, A.}, {Gastineau,
  M.}, and {Manche, H.}]{Laskar2011}
{Laskar, J.}; {Fienga, A}.; {Gastineau, M}.; {Manche, H}.
\newblock La2010: A new orbital solution for the long-term motion of the Earth.
\newblock {\em Astron. Astrophys.} {\bf 2011}, {\em 532},~A89.
\newblock
  {https://doi.org/10.1051/0004-6361/201116836}.

\bibitem[{Laskar, J.} {et~al.}(2004){Laskar, J.}, {Robutel, P.}, {Joutel,
  F.}, {Gastineau, M.}, {Correia, A. C. M.}, and {Levrard, B.}]{Laskar2004}
{Laskar, J.}; {Robutel, P}.; {Joutel, F}.; {Gastineau, M.}; {Correia, A.C.M.}; {Levrard, B}.
\newblock A long-term numerical solution for the insolation quantities
  of~the~Earth.
\newblock {\em Astron. Astrophys.} {\bf 2004}, {\em 428},~261--285.
\newblock
  {https://doi.org/10.1051/0004-6361:20041335}.

\bibitem[Westerhold {et~al.}(2020)Westerhold, Marwan, Drury, Liebrand,
  Agnini, Anagnostou, Barnet, Bohaty, De~Vleeschouwer, Florindo, Frederichs,
  Hodell, Holbourn, Kroon, Lauretano, Littler, Lourens, Lyle, P{\"a}like,
  R{\"o}hl, Tian, Wilkens, Wilson, and Zachos]{Westerhold2020}
Westerhold, T.; Marwan, N.; Drury, A.J.; Liebrand, D.; Agnini, C.; Anagnostou,
  E.; Barnet, J.S.K.; Bohaty, S.M.; De~Vleeschouwer, D.; Florindo, F.;  et~al.
\newblock An astronomically dated record of Earth{\textquoteright}s climate and
  its predictability over the last 66 million years.
\newblock {\em Science} {\bf 2020}, {\em 369},~1383--1387.
\newblock
  {https://doi.org/10.1126/science.aba6853}.

\bibitem[Kraemer {et~al.}(2022)Kraemer, Gelbrecht, Pavithran, Sujith, and
  Marwan]{Kraemer2022}
Kraemer, K.H.; Gelbrecht, M.; Pavithran, I.; Sujith, R.I.; Marwan, N.
\newblock Optimal state space reconstruction via Monte Carlo decision tree
  search.
\newblock {\em Nonlinear Dyn.} {\bf 2022}, {\em 108},~1525--1545.
\newblock
  {https://doi.org/10.1007/s11071-022-07280-2}.

\bibitem[Hegger and Kantz(1999)]{hegger1999}
Hegger, R.; Kantz, H.
\newblock Improved false nearest neighbor method to detect determinism in time
  series data.
\newblock {\em Phys. Rev. E} {\bf 1999}, {\em 60},~4970--4973.
\newblock
  {https://doi.org/10.1103/PhysRevE.60.4970}.

\bibitem[Pecora {et~al.}(2007)Pecora, Moniz, Nichols, and
  Carroll]{pecora2007}
Pecora, L.M.; Moniz, L.; Nichols, J.; Carroll, T.L.
\newblock A unified approach to attractor reconstruction.
\newblock {\em Chaos Interdiscip. J. Nonlinear Sci.} {\bf
  2007}, {\em 17},~013110.
\newblock
  {https://doi.org/10.1063/1.2430294}.

\bibitem[Datseris(2018)]{datseris2018}
Datseris, G.
\newblock DynamicalSystems.jl: A Julia software library for chaos and nonlinear
  dynamics.
\newblock {\em J. Open Source Softw.} {\bf 2018}, {\em 3},~598.
\newblock
  {https://doi.org/10.21105/joss.00598}.

\bibitem[Rackauckas and Nie(2017)]{rackauckas2017}
Rackauckas, C.; Nie, Q.
\newblock Differentialequations.jl--a performant and feature-rich ecosystem for
  solving differential equations in julia.
\newblock {\em J. Open Res. Softw.} {\bf 2017}, {\em 5}, 15.

\bibitem[Zhang {et~al.}(2022)Zhang, Widmann, and Barreira]{Zhang2022}
\textls[-15]{Zhang, S.; Widmann, D.; Barreira, D.S}.
\newblock {JuliaOptimalTransport/OptimalTransport.jl: v0.3.19}%MDPI: Please provide more information about the type of the article, such as book (please provide the name of the publisher and location of it), online resource (please provide the URL of the website and the date it was accessed (day, month, year)), or journal article (please provide the name of the journal, the year and volume it published, and the page number).
. {2022}.
\newblock
   \hl{Available online:} %MDPI: we replace the correct web link, please confirm
 \url{https://zenodo.org/record/5920148\#.Y3SbbORByUk} \hl{(accessed on)} %MDPI: Please add accessed date.

\bibitem[Lorenz(1963)]{lorenz1963}
Lorenz, E.N.
\newblock Deterministic Nonperiodic Flow.
\newblock {\em J. Atmos. Sci.} {\bf 1963}, {\em
  20},~130--141.
\newblock
  {https://doi.org/10.1175/1520-0469(1963)020<0130:DNF>2.0.CO;2}.

\bibitem[Rössler(1976)]{roessler1976}
Rössler, O.
\newblock An equation for continuous chaos.
\newblock {\em Phys. Lett. A} {\bf 1976}, {\em 57},~397 -- 398.
\newblock
  {https://doi.org/10.1016/0375-9601(76)90101-8}.

\bibitem[GB()]{GB}
National Grid, Frequency Data (2014–2018).
\newblock  \hl{Available online}%MDPI: please add the website link.
:  (accessed on 25 July 2022).

\bibitem[Haehne {et~al.}(2018)Haehne, Schottler, Waechter, Peinke, and
  Kamps]{haehne2018footprint}
Haehne, H.; Schottler, J.; Waechter, M.; Peinke, J.; Kamps, O.
\newblock The footprint of atmospheric turbulence in power grid frequency
  measurements.
\newblock {\em EPL (Europhys. Lett.)} {\bf 2018}, {\em 121},~30001.

\bibitem[Fraser and Swinney(1986)]{fraser1986}
Fraser, A.M.; Swinney, H.L.
\newblock Independent coordinates for strange attractors from mutual
  information.
\newblock {\em Phys. Rev. A} {\bf 1986}, {\em 33},~1134--1140.
\newblock
  {https://doi.org/10.1103/PhysRevA.33.1134}.

\end{thebibliography}

%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{adjustwidth}
\end{document}